---
title: "Cholorphyll a & c2"
author: "HM Putnam, AS Huffmyer"
editied: "Florence Fields"
date: "2025-01-29"
project:ENCORE
output: html_document
---

This script calculates chlorphyll a and c2 content in Mdec and Dlab samples from the Priming experiment conducted in Bermuda 2024 and is an edited version of scipts from Jill and E5 edited by Hollie and Arianna https://github.com/daniellembecker/A.pul_Heatwave/blob/master/heatwave/timeseries/timepoint1/1_scripts/1_chlorophyll.Rmd

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

setwd("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024")

```


## Install packages if you dont already have them
```{r}

if (!require("tidyverse")) install.packages("tidyverse")
if (!require("plotrix")) install.packages("plotrix")
if (!require("dplyr")) install.packages("dplyr")
if (!require("Rmisc")) install.packages("Rmisc")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("gridExtra")) install.packages("gridExtra")
if (!require("formattable")) install.packages("formattable")
if (!require("data.table")) install.packages("data.table")
if (!require("htmltools")) install.packages("htmltools")
if (!require("webshot")) install.packages("webshot")

```

## Load packages
```{r}

library(plyr)
library(dplyr)
library(tidyverse)
library(plotrix)
library(Rmisc)
library(ggplot2)
library(gridExtra)
library(data.table)
library(htmltools)
library(webshot)
library(lme4)
library(MuMIn)
library(cowplot)
library(corrplot)
library(lmerTest)
library(scales)
library(formattable)
library(tidyr)
library(purrr)
library(readr)
library(reshape2)
library(car)

```

## Importing data and merging into one large dataframe
```{r}

# Define function to read in chl data
read_chl <- function(file) {
  chl_data <- read_csv(file, skip = 24, n_max = 24)%>%
    select(-1)%>%
    magrittr::set_colnames(c("row", 1:12, "wavelength"))%>%
    fill(row)%>%
    gather("col", "absorbance", -wavelength, -row)%>%
    unite("well", c(row, col), sep = "")
}

# List chlorophyll data files
chl_path <- "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/Chlorophyll/"                               # Path to chlorophyll data directory
all_chl_files <- list.files(path = chl_path, pattern = "*.csv")          # List all files in directory
chl_platemaps <- list.files(path = chl_path, pattern = "platemap")       # List platemap files
chl_data_files <- setdiff(all_chl_files, chl_platemaps)                  # List absorbance data files

# Read in all files into tibble
df <- tibble(file = chl_data_files) %>%
  mutate(platemap = map(file, ~ read_csv(paste0(chl_path, tools::file_path_sans_ext(.), "_platemap.csv"))),
         chl_data = map(file, ~ read_chl(paste0(chl_path, .))))


# Print the resulting tibble
print(df)



# Merge platemap and data for each plate
df <- df %>%
  mutate(merged = map2(platemap, chl_data, ~ right_join(.x, .y)))

```

# Calculate chlorophyll concentrations
```{r}
# average all technical replicates for each plate/sample/wavelength, including all acetone blanks together (per plate)

df1 <- df %>%
  unnest(merged) %>% # Unnest the merged column to access individual rows
  group_by(file, Fragment_ID, wavelength) %>% # Group by plate, sample, and wavelength
  summarize(avg_absorbance = mean(absorbance, na.rm = TRUE)) %>% # Calculate the average absorbance
  filter(!is.na(Fragment_ID)) %>%
  spread(wavelength,avg_absorbance )


# get the acetone blank 750 absorbace for each file (i.e., plate), and subtract from 630 and 663 values for each sample
df1 <- df1 %>%
  group_by(file) %>%
  mutate(blank750 = `Chl:750`[Fragment_ID== "BLK"]) %>%
  ungroup() %>%
  mutate(adj630 = `Chl:630` - blank750,
         adj663 = `Chl:663` - blank750)

# calculate chla and chlc2 values based on equations from Jeffrey and Humphrey 1975
# units µg/ml
#path length adjustment = 0.6 

df1 <- df1 %>%
  mutate(chla.ug.ml = (11.43 * adj663)/0.6 - (0.64 * adj630)/0.6,
        chlc2.ug.ml = (27.09 * adj630)/0.6 - (3.63 * adj663)/0.6)
```

# Normalize to surface area
```{r}
# Load homogenate volume
homog.vol <- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/Airbrushing_data.csv")%>%
  select(Species, Fragment_ID, Homogenate.Vol)
chl <- full_join(df1, homog.vol)

# Load surface area
sa <- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/Surface_Area/Surface_Area.csv")%>%
rename(Fragment_ID = Colony_id)
chl1 <- full_join(chl, sa)

# Multiply chlorophyll by the homogenate volume and divide by surface area
chl1 <- chl1 %>%
  mutate(chla.ug.cm2 = chla.ug.ml * Homogenate.Vol / surface.area.cm2,
         chlc2.ug.cm2 = chlc2.ug.ml * Homogenate.Vol / surface.area.cm2)

# write chlorophyll data to file and remove blanks and NAs
chl1<- chl1%>%
  select(Fragment_ID, Species, Sample_date, Tank, chla.ug.cm2, chlc2.ug.cm2)%>%
  mutate(timepoint= case_when(                   #Adding the timepoint colunm corelating the timepoints with the sample dates 
    Sample_date == "20240620" ~ "TP0",
    Sample_date == "20240709" ~ "TP1",
    Sample_date == "20240723" ~ "TP2",
    Sample_date == "20240806" ~ "TP3",
    Sample_date == "20240815" ~ "TP4"
    ),
    Treatment = case_when(
      Tank == "2" ~ "Low_pH",
    Tank == "3" ~ "Low_pH",
    Tank == "4" ~ "Control",
    Tank == "5" ~ "Control",
    Tank == "6" ~ "High_Temp",
    Tank == "7" ~ "High_Temp",
    Tank == "acclimation" ~ "Acclimation"
    ),
    Colony = case_when(
      grepl("^DL-1($|-)", Fragment_ID) ~ 1,
      grepl("^DL-2($|-)", Fragment_ID) ~ 2,
      grepl("^DL-3($|-)", Fragment_ID) ~ 3,
      grepl("^DL-4($|-)", Fragment_ID) ~ 4,
      grepl("^MD-1($|-)", Fragment_ID) ~ 1,
      grepl("^MD-2($|-)", Fragment_ID) ~ 2,
      grepl("^MD-4($|-)", Fragment_ID) ~ 4,
      grepl("^MD-5($|-)", Fragment_ID) ~ 3,
      TRUE ~ NA_integer_
                     ))%>%
  filter(!is.na(chla.ug.cm2))
```


#Adding Secondary treatment info using the coral fragment metadata file
```{r}

frag_meta <- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/Coral_Fragment_Metadata.csv") %>%
  mutate(Sample.ID = gsub("_", "-", Sample.ID))%>%
  write_csv(frag_meta, path = "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/Coral_Fragment_Metadata1.csv" )

frag_meta <- frag_meta%>%
  select(Sample.ID, Inital.Treament, Secondary.Treatment)%>%
  rename(Inital.Treatment = Inital.Treament)%>%
  rename(Fragment_ID = Sample.ID)

chl2 <- left_join (chl1, frag_meta, by = "Fragment_ID")


chl2<- chl2%>%
  select(Fragment_ID, Species, Sample_date, Tank, chla.ug.cm2, chlc2.ug.cm2, timepoint, Treatment, Secondary.Treatment, Colony)%>%
rename(Inital.Treatment = Treatment)

  
# Print the updated data frame
print(chl2)

write_csv(chl2, path = "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/chlorophyll/chlorophyll.csv")

#Removing timepoints 0 and 1
chl3 <- chl2 %>%
  filter(timepoint != "TP0",
         timepoint !="TP1")

```

Checking for outliers 
```{r}

#Check and remove outliers in colunm chla.ug.cm2

# Calculate the IQR
Q1 <- quantile(chl3$chla.ug.cm2, 0.25)
Q3 <- quantile(chl3$chla.ug.cm2, 0.75)
IQR <- Q3 - Q1

# Identify outliers
outliers <- chl3$chla.ug.cm2[(chl3$chla.ug.cm2 < (Q1 - 1.5 * IQR)) | (chl3$chla.ug.cm2 > (Q3 + 1.5 * IQR))]
print(outliers)

chl3 <- chl3 %>%
  filter(!(chla.ug.cm2 %in% outliers))

#Check and remove outliers in colunm chlc2.ug.cm2

Q1 <- quantile(chl3$chlc2.ug.cm2, 0.25)
Q3 <- quantile(chl3$chlc2.ug.cm2, 0.75)
IQR <- Q3 - Q1

outliers <- chl3$chlc2.ug.cm2[(chl3$chlc2.ug.cm2 < (Q1 - 1.5 * IQR)) | (chl3$chlc2.ug.cm2 > (Q3 + 1.5 * IQR))]
print(outliers)

chl3 <- chl3 %>%
  filter(!(chlc2.ug.cm2 %in% outliers))

```


#Standard error calculations
```{r}
#For visualization purposes there will be 2 seperate datasets, one with just initial treatment, another with secondary treatment 

# Calculate the standard error for each species and timepoint with initial treatment
df_summary <- chl3 %>%
  group_by(Species, timepoint, Inital.Treatment,Tank, Sample_date) %>%
  summarize(
    chla_mean = mean(chla.ug.cm2),
    chla_se = sd(chla.ug.cm2) / sqrt(n()),
    chlc2_mean = mean(chlc2.ug.cm2),
    chlc2_se = sd(chlc2.ug.cm2) / sqrt(n())
  )
# Calculate the standard error for each species and timepoint with initial and secondary treatment, without TP0 and acclimation

df_summary2 <- chl3 %>%
 # filter(!grepl("NONE", Secondary.Treatment))%>%
  group_by(Species, timepoint, Inital.Treatment, Secondary.Treatment,Tank,Sample_date) %>%
  summarize(
    chla_mean = mean(chla.ug.cm2),
    chla_se = sd(chla.ug.cm2) / sqrt(n()),
    chlc2_mean = mean(chlc2.ug.cm2),
    chlc2_se = sd(chlc2.ug.cm2) / sqrt(n())
  )

print(df_summary2)

#Only Madracis decactis
df_summary3 <- df_summary2 %>%
  filter(!grepl("Diploria labrinthiformis", Species))

#Only Diploria labrinthiformis
df_summary4 <- df_summary2 %>%
  filter(!grepl("Madracis decactis", Species))

#Averaging treatment groups in Timepoint 2 and 3


df_summary2_avg <- df_summary2 %>%
  mutate(Tank_Group = case_when(
    Tank %in% c(2, 3) ~ "Tank_2_3",
    Tank %in% c(4, 5) ~ "Tank_4_5",
    Tank %in% c(6, 7) ~ "Tank_6_7",
    TRUE ~ as.character(Tank)
  )) %>%
  group_by(Species, timepoint, Inital.Treatment, Secondary.Treatment, Tank_Group) %>%
  summarize(
    chla_mean = mean(chla_mean), 
    chla_se = sqrt(sum(chla_se^2)/n()),
    chlc2_mean = mean(chlc2_mean), 
    chlc2_se = sqrt(sum(chlc2_se^2)/n()),
    .groups = "drop"
  )
write_csv(df_summary2_avg, path = "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/chlorophyll/chlorophylldfsummary2_avg.csv")
```

#Including a column that includes the treatment stage (Priming, recovery and Stress)
```{r}
df_summary2_avg<- df_summary2_avg%>%
  mutate(TreatmentStage= case_when( timepoint == "TP2" ~ "Priming",
                                    timepoint == "TP3" ~ "Recovery",
                                    timepoint == "TP4" ~ "Stress"))

```

```{r}
# Generate the plot Chla using the means

df_summary2_avg$SecondaryColor <- ifelse(df_summary2_avg$timepoint == "TP4", df_summary2_avg$Secondary.Treatment, NA)



ggplot(df_summary2_avg, aes(x = Inital.Treatment, y = chla_mean, color = Inital.Treatment, fill = SecondaryColor)) +
  geom_errorbar(aes(ymin = chla_mean - chla_se, ymax = chla_mean + chla_se), 
                width = 0.2, position = position_dodge(width = 0.75)) +
  geom_point(position = position_dodge(width = 0.75), shape = 21, size = 3) +
  facet_grid(Species ~ TreatmentStage) +
  theme_classic()+
  scale_color_manual(
    name = "Initial Treatment",
    values = c("Control" = "lightblue", "High_Temp" = "red", "Low_pH" = "purple")) +
  scale_fill_manual(
    name = "Secondary Color",
    values = c("Control" = "blue", "High_Temp" = "darkred", "NA" = "grey")) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        panel.spacing = unit(0.3, "lines"),
     panel.border = element_rect(color = "black", fill = NA)  )+
  labs(x = "Initial Treatment", y = expression("Average Chlorophyll a (µg/cm"^2*")"))+
ggsave("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/chlorophyll/Chla3.pdf", device = "pdf", width = 8, height = 6)


```

#PLOT chlc2 
```{r}
#Generate the plot Chlc2 using means
ggplot(df_summary2_avg, aes(x = Inital.Treatment, y = chlc2_mean, color = Inital.Treatment, fill = SecondaryColor)) +
  geom_errorbar(aes(ymin = chlc2_mean - chlc2_se, ymax = chlc2_mean + chlc2_se), 
                width = 0.2, position = position_dodge(width = 0.75)) +
  geom_point(position = position_dodge(width = 0.75), shape = 21, size = 3) +
  facet_grid(Species ~ TreatmentStage) +
  theme_classic()+
  scale_color_manual(
    name = "Initial Treatment",
    values = c("Control" = "lightblue", "High_Temp" = "red", "Low_pH" = "purple")) +
  scale_fill_manual(
    name = "Secondary Color",
    values = c("Control" = "blue", "High_Temp" = "darkred", "NA" = "grey")) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        panel.spacing = unit(0.3, "lines"),
     panel.border = element_rect(color = "black", fill = NA)  )+
  labs(x = "Initial Treatment", y = expression("Average Chlorophyll C2 (µg/cm"^2*")"))+
ggsave("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/chlorophyll/Chlc2-3.pdf", device = "pdf", width = 8, height = 6)

```
TP2D<-TP2%>%filter(Species != "Madracis decactis")




# Statistical Test within timepoints 
#TP2 = Priming
```{r}

##Separating time points to test individually 

TP2<- chl2%>%
  filter(timepoint != "TP0",
         timepoint !="TP1",
         timepoint !="TP3",
         timepoint !="TP4")

## Test for differences in TP2 for chla and chlc2, treatment and species with a two-way ANOVA

model <- aov(chla.ug.cm2 ~ Inital.Treatment*Species, data = TP2)
summary(model)

#Df Sum Sq Mean Sq F value Pr(>F)
#Inital.Treatment          2  1.132  0.5659   1.269  0.292
#Species                   1  0.113  0.1129   0.253  0.618
#Inital.Treatment:Species  2  0.022  0.0108   0.024  0.976
#Residuals                41 18.289  0.4461    

#There is no significant effect of initial treatment on chla or between species at during priming

TukeyHSD(model) # See if there is any differences between specific treatments 
plot(model, 1) # Check ANOVA assumptions of normality 

leveneTest(chla.ug.cm2 ~ Inital.Treatment, data = TP2) # Check ANOVA assumptions of homogenity of variance  

#Df F value Pr(>F)
#group  2   0.118  0.889
#      44   

interaction.plot(TP2$Inital.Treatment, TP2$Species, TP2$chla.ug.cm2)
#The interaction plot shows no interaction effect

model <- aov(chlc2.ug.cm2 ~ Inital.Treatment*Species, data = TP2)
summary(model)

#                         Df Sum Sq Mean Sq F value Pr(>F)
#Inital.Treatment          2 0.0542 0.02711   0.649  0.528
#Species                   1 0.0065 0.00651   0.156  0.695
#Inital.Treatment:Species  2 0.0089 0.00443   0.106  0.900
#Residuals                41 1.7117 0.04175  

#No significant effect of initial treatment between species on chlc2 during priming

TukeyHSD(model) # See if there is any differences between specific treatments 

plot(model, 1) # Check ANOVA assumptions of normality 

leveneTest(chlc2.ug.cm2 ~ Inital.Treatment, data = TP2) # Check ANOVA assumptions of homogenity of variance   

#Df F value Pr(>F)
#group  2  0.9097 0.4101
#      44 

interaction.plot(TP2$Inital.Treatment, TP2$Species, TP2$chlc2.ug.cm2)
#The interaction plot shows an interaction effect in response to low pH, meaning different species respond differently to low pH
```

#TP3 = Recovery
```{r}
TP3<- chl2%>%
  filter(timepoint != "TP0",
         timepoint !="TP1",
         timepoint !="TP2",
         timepoint !="TP4")

model <- aov(chla.ug.cm2 ~ Inital.Treatment*Species, data = TP3)
summary(model)

#                         Df Sum Sq Mean Sq F value Pr(>F)
#Inital.Treatment          2  0.198 0.09914   0.474  0.626
#Species                   1  0.031 0.03051   0.146  0.704
#Inital.Treatment:Species  2  0.589 0.29473   1.410  0.255
#Residuals                43  8.988 0.20902     

#There is no significant effect of initial treatment on chla or between species at during Recovery

TukeyHSD(model) 
plot(model, 1)  

leveneTest(chla.ug.cm2 ~ Inital.Treatment, data = TP3)

#Df F value Pr(>F)
#group  2  1.2151  0.306
#      46 

interaction.plot(TP3$Inital.Treatment, TP3$Species, TP3$chla.ug.cm2)
#lines cross or diverge, suggesting an interaction effect, meaning different the two species respond differently to the treatments.


model <- aov(chlc2.ug.cm2 ~ Inital.Treatment*Species, data = TP3)
summary(model)

#                        Df Sum Sq Mean Sq F value Pr(>F)
#Inital.Treatment          2 0.0517 0.02585   0.982  0.383
#Species                   1 0.0065 0.00646   0.245  0.623
#Inital.Treatment:Species  2 0.0848 0.04239   1.610  0.212
#Residuals                43 1.1319 0.02632 

#Significant effect of initial treatment between species on chlc2 during priming

TukeyHSD(model) # See if there is any differences between specific treatments 

plot(model, 1) # Check ANOVA assumptions of normality 

leveneTest(chlc2.ug.cm2 ~ Inital.Treatment, data = TP3) 

#     Df F value Pr(>F)
#group  2  0.4627 0.6325
#      46

interaction.plot(TP3$Inital.Treatment, TP3$Species, TP3$chlc2.ug.cm2)
#lines cross or diverge, suggesting an interaction effect, meaning different the two species respond differently to the treatments.

```

#TP4 = Stress
```{r}
TP4<- chl2%>%
  filter(timepoint != "TP0",
         timepoint !="TP1",
         timepoint !="TP2",
         timepoint !="TP3")

model <- aov(chla.ug.cm2 ~ Secondary.Treatment*Species, data = TP4)
summary(model)
#                            Df Sum Sq Mean Sq F value   Pr(>F)    
#Secondary.Treatment          1  1.538  1.5378  12.121 0.000749 ***
#Species                      1  1.398  1.3984  11.022 0.001270 ** 
#Secondary.Treatment:Species  1  0.166  0.1656   1.305 0.256066    
#Residuals                   97 12.307  0.1269   

#Statistical effect seen between chla and species

TukeyHSD(model)  
plot(model, 1) 

leveneTest(chla.ug.cm2 ~ Secondary.Treatment, data = TP4) 
#Df F value  Pr(>F)  
#group  1  0.4505 0.5037
#      99               

interaction.plot(TP4$Secondary.Treatment, TP4$Species, TP4$chla.ug.cm2)


model <- aov(chlc2.ug.cm2 ~ Secondary.Treatment*Species, data = TP4)
summary(model)

#                            Df Sum Sq Mean Sq F value   Pr(>F)    
#Secondary.Treatment          1 0.0662 0.06618   4.658   0.0334 *  
#Species                      1 0.2663 0.26631  18.745 3.64e-05 ***
#Secondary.Treatment:Species  1 0.0009 0.00089   0.063   0.8028    
#Residuals                   97 1.3781 0.01421  


TukeyHSD(model) # See if there is any differences between specific treatments 
plot(model, 1) # Check ANOVA assumptions of normality 
leveneTest(chlc2.ug.cm2 ~ Secondary.Treatment, data = TP4) 

# Df F value Pr(>F)
#group  1  0.0052 0.9426
#      99 

interaction.plot(TP4$Inital.Treatment, TP4$Species, TP4$chlc2.ug.cm2)
#Interaction effect
```

#Between Timepoints
```{r}
model <- aov(chla.ug.cm2~Inital.Treatment*Secondary.Treatment*Species, data = chl3)
summary(model)
```

```{r}
TukeyHSD(model)  
plot(model, 1) 
interaction.plot(chl3$Secondary.Treatment, chl3$Species, chl3$chlc2.ug.cm2)

```
```{r}
model <- aov(chlc2.ug.cm2~Inital.Treatment*Secondary.Treatment*Species, data = chl3)
summary(model)

```
```{r}
TukeyHSD(model)  
plot(model, 1) 
interaction.plot(chl3$Secondary.Treatment, chl3$Species, chl3$chla.ug.cm2)
```