---
title: "PAM data analysis"
author: "Florence Fields"
date: "2024-06-12"
output: html_document
---

This code analyzes the pulse amplitude fluometry (PAM) data that was collected from from adult Mdec and Dlab in Bermuda 2024 as part of the ENCORE priming experiments. 

## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r}
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(car)
library(emmeans)
library(multcomp)
library(knitr)
library(kableExtra)
library(lme4)
library(GGally)
library(lmerTest)
library(cowplot)
library(Hmisc)
library(ggstatsplot)
library(dplyr)
```

## First, I am going to analyze the data from the larval samples. 

## Load data 
```{r}
pam_data <- read.csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/Pam/All_Pam_Data.csv", na.strings=c("","NA"))
head(pam_data)
tail(pam_data)
pam_data$Colony <- ifelse(pam_data$Colony == 5, 3, pam_data$Colony)
```

## Manipulate data frame 

```{r}

pam_data <- pam_data %>%
  dplyr::mutate(Treatment = case_when(
    Tank %in% c(6, 7) ~ "High Heat",
    Tank %in% c(2, 3) ~ "Low pH",
    Tank %in% c(4, 5) ~ "Control",
    TRUE ~ Treatment  # Keep other values unchanged
  ))
```

rename colounm heads 
```{r}
colnames(pam_data)[colnames(pam_data) == "Fo.I."] <- "Fo.I"
colnames(pam_data)[colnames(pam_data) == "Fo.II."] <- "Fo.II"
colnames(pam_data)[colnames(pam_data) == "Fo.III."] <- "Fo.III"
colnames(pam_data)[colnames(pam_data) == "Fm.I."] <- "Fm.I"
colnames(pam_data)[colnames(pam_data) == "Fm.II."] <- "Fm.II"
colnames(pam_data)[colnames(pam_data) == "Fm.III."] <- "Fm.III"
colnames(pam_data)[colnames(pam_data) == "Y.I."] <- "Fv.Fm.I"
colnames(pam_data)[colnames(pam_data) == "Y.II."] <- "Fv.Fm.II"
colnames(pam_data)[colnames(pam_data) == "Y.III."] <- "Fv.Fm.III"
colnames(pam_data)[colnames(pam_data) == "Fo..avg."] <- "Fo.avg"
colnames(pam_data)[colnames(pam_data) == "Fm..avg."] <- "Fm.avg"
colnames(pam_data)[colnames(pam_data) == "Y.avg."] <- "FvFm_avg"

as.character(pam_data$DateMeasured)

pam_data<- pam_data %>%
  mutate(Species = ifelse(Species == "DL", "Diploria labyrinthiformis", Species))
pam_data<- pam_data %>%
  mutate(Species = ifelse(Species == "MD", "Madracis decactis", Species))
pam_data<- pam_data %>%
  mutate(Tank = ifelse(Tank == "8", "Acclimation", Tank))
```

#Standard error calculations
```{r}

# Ensure there are no NA values and check the structure of the data
pam_data <- pam_data %>%
  filter(!is.na(FvFm_avg))

# Calculate the average and standard error for each species and timepoint
pam_summary <- pam_data %>%
  group_by(Species, DateMeasured, Treatment, Tank) %>%
  dplyr::summarize(
    Fv.Fm_Mean = mean(FvFm_avg, na.rm = TRUE),
    Fv.Fm_SE = sd(FvFm_avg, na.rm = TRUE) / sqrt(dplyr::n()))

# Print the summary data frame
print(pam_summary)

# Save the dataset as a CSV file
write.csv(pam_summary, file = "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/pam/pam_summary.csv")


# Print the first few rows of the filtered data for debugging
print(head(pam_summary))


```

#Creating a ssecondary treatment colunm
```{r}

pam_data2 <- pam_data%>%
  mutate(secondary_treatment= case_when(                    
    DateMeasured %in% c("20240807", "20240809", "20240815") & Tank %in% c(2, 4, 6) ~ "High Heat",
    DateMeasured %in% c("20240807", "20240809", "20240815") & Tank %in% c(3, 5, 7) ~ "Control",
      TRUE ~ NA_character_  # Assign NA to all other rows
    )
  )
pam_data2$DateMeasured <- as.Date(as.character(pam_data2$DateMeasured), format = "%Y%m%d")


pam_data3<- pam_data2 %>%
  dplyr::select(DateMeasured, Species, Treatment, Tank, FvFm_avg, secondary_treatment)%>%
  mutate(Treatmentstage = case_when(
    DateMeasured == "2024-08-07" ~ "Stress",
    DateMeasured == "2024-08-09" ~ "Stress",
    DateMeasured == "2024-08-15" ~ "Stress",
    DateMeasured == "2024-07-02" ~ "Acclimation",
    DateMeasured == "2024-07-19" ~ "Priming",
    DateMeasured == "2024-07-23" ~ "Priming",
    DateMeasured == "2024-07-26" ~ "Recovery",
    DateMeasured == "2024-07-30" ~ "Recovery",
    DateMeasured == "2024-07-05" ~ "Acclimation",
    TRUE ~ as.character(DateMeasured)
  ))

#remove outliers
# Calculate the IQR
Q1 <- quantile(pam_data3$FvFm_avg, 0.25)
Q3 <- quantile(pam_data3$FvFm_avg, 0.75)
IQR <- Q3 - Q1

# Identify outliers
outliers <- pam_data3$FvFm_avg[(pam_data3$FvFm_avg < (Q1 - 1.5 * IQR)) | (pam_data3$FvFm_avg > (Q3 + 1.5 * IQR))]
print(outliers)

pam_data4 <- pam_data3 %>%
  filter(!(FvFm_avg %in% outliers))

# Calculate the average and standard error 
pam_summary2 <- pam_data3 %>%
  group_by(Species, DateMeasured, Treatment, secondary_treatment, Treatmentstage) %>%
  dplyr::summarize(
    Fv.Fm_Mean = mean(FvFm_avg, na.rm = TRUE),
    Fv.Fm_SE = sd(FvFm_avg, na.rm = TRUE) / sqrt(dplyr::n()))

pam_summary3 <- pam_data4 %>%
  group_by(Species, DateMeasured, Treatment, secondary_treatment, Treatmentstage) %>%
  dplyr::summarize(
    Fv.Fm_Mean = mean(FvFm_avg, na.rm = TRUE),
    Fv.Fm_SE = sd(FvFm_avg, na.rm = TRUE) / sqrt(dplyr::n()))

#Removing Acclimation rows with outliers
pam_summary4 <- pam_summary2 %>%
  filter(Treatment != "Acclimation")


pam_summary5 <- pam_summary3 %>%
  filter(Treatment != "Acclimation")



```



Plot based off of pam_data3
```{r}

#Generate the plot using means

ggplot(pam_summary4, aes(as.character(DateMeasured), Fv.Fm_Mean, color = Treatment, fill = secondary_treatment)) +
  geom_errorbar(aes(ymin = Fv.Fm_Mean - Fv.Fm_SE, ymax = Fv.Fm_Mean + Fv.Fm_SE), 
                width = 0.3, position = position_dodge(width = 0.75), linewidth = 0.5) +
  geom_point(aes(shape = Treatment), position = position_dodge(width = 0.75), size = 3)+
  facet_grid(Species ~ Treatmentstage, scales = "free_x")+
  theme_classic()+
  scale_color_manual(values = c("Control" = "lightblue", "High Heat" = "red", "Low pH" = "purple")) +
  scale_fill_manual(values = c("Control" = "blue", "High Heat" = "darkred", "NA" = "grey"))+
  scale_shape_manual(values = c("Control" = 16, "High Heat" = 17, "Low pH" = 18))+
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        panel.spacing = unit(0.3, "lines"),
     panel.border = element_rect(color = "black", fill = NA)  )+
  labs(x = "Date", y = "Fv/Fm")
ggsave("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/pam/pam_point.pdf", device = "pdf", width = 8, height = 6)



pam_summary4 <- pam_summary4[pam_summary4$DateMeasured != as.Date("2024-08-09"), ]

ggplot(pam_summary4, aes(x = as.character(DateMeasured), y = Fv.Fm_Mean, 
                         color = ifelse(Treatmentstage == "Stress", secondary_treatment, "black"), 
                         fill = ifelse(Treatmentstage == "Stress", secondary_treatment, "black"),
                         group = Treatment)) + 
  geom_errorbar(aes(ymin = Fv.Fm_Mean - Fv.Fm_SE, ymax = Fv.Fm_Mean + Fv.Fm_SE), 
                width = 0.3, position = position_dodge(width = 0.75), linewidth = 0.5) + 
  geom_point(aes(shape = Treatment), position = position_dodge(width = 0.75), size = 3) + 
  facet_grid(Species ~ Treatmentstage, scales = "free_x") + 
  theme_classic() + 
  scale_color_manual(name = "Secondary Treatment", values = c("Control" = "blue", "High Heat" = "darkred", "Low pH" = "purple", "NA" = "black")) + 
  scale_fill_manual(name = "Secondary Treatment",
                    values = c("Control" = "blue", "High Heat" = "darkred", "Low pH" = "purple", "NA" = "black")) + 
  scale_shape_manual(name = "Primary Treatment", values = c("Control" = 16, "High Heat" = 17, "Low pH" = 18)) + 
  theme(axis.text.x = element_text(angle = 35, hjust = 1), 
        panel.spacing = unit(0.3, "lines"), 
        panel.border = element_rect(color = "black", fill = NA)) + 
  labs(x = "Date", y = "Fv/Fm")
ggsave("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/pam/pam_point_final2.pdf", device = "pdf", width = 8, height = 6)
```



Do stats on Fv/Fm by treatment, species and date on Dlab
```{r}
DL <- pam_data3 %>%
  filter(!grepl("Madracis decactis", Species))

#DL STATS

model <- aov(FvFm_avg ~ DateMeasured*Treatment*Treatmentstage, data = DL)
summary(model)
#Shows significant differences between species, treatment, treatment-species, date-treatement and between date-species-treatment. 
Dlab
                                       Df Sum Sq Mean Sq F value   Pr(>F)    
DateMeasured                            1 0.0569 0.05690  35.551 4.51e-09 ***
Treatment                               3 0.1408 0.04694  29.326  < 2e-16 ***
Treatmentstage                          2 0.0605 0.03027  18.911 1.16e-08 ***
DateMeasured:Treatment                  3 0.0425 0.01415   8.844 9.91e-06 ***
DateMeasured:Treatmentstage             2 0.0115 0.00576   3.600    0.028 *  
Treatment:Treatmentstage                4 0.0055 0.00136   0.852    0.493    
DateMeasured:Treatment:Treatmentstage   3 0.0080 0.00267   1.665    0.173    
Residuals                             537 0.8595 0.00160                     
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

# View post-hoc differences 
emm = emmeans(model, ~ Treatment*DateMeasured)
cld(emm, Letters=c(LETTERS)) #letter display
pairs(emm)
TukeyHSD(model)
```
Dlab

```{R}
MD <- pam_data3 %>%
  filter(!grepl("Diploria labyrinthiformis", Species))

#MD STATS

model <- aov(FvFm_avg ~ DateMeasured*Treatment*Treatmentstage, data = MD)
summary(model)
#Shows significant differences between species, treatment, treatment-species, date-treatement and between date-species-treatment. 
Mdec
                                       Df Sum Sq Mean Sq F value Pr(>F)    
DateMeasured                            1 0.1964 0.19641 117.407 <2e-16 ***
Treatment                               3 0.5174 0.17246 103.095 <2e-16 ***
Treatmentstage                          2 0.0005 0.00025   0.148 0.8621    
DateMeasured:Treatment                  3 0.0133 0.00443   2.645 0.0488 *  
DateMeasured:Treatmentstage             2 0.0065 0.00327   1.958 0.1425    
Treatment:Treatmentstage                4 0.0073 0.00182   1.086 0.3632    
DateMeasured:Treatment:Treatmentstage   3 0.0182 0.00607   3.626 0.0132 *  
Residuals                             411 0.6875 0.00167                   
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

# View post-hoc differences 
emm = emmeans(model, ~ Intial_treatment*Date)
cld(emm, Letters=c(LETTERS)) #letter display
pairs(emm)

TukeyHSD(model)
```
Mdec

```{r}
pam_data4$SecondaryColor <- ifelse(pam_data3$Date %in% c("2024-08-07", "2024-08-09", "2024-08-15"), pam_data4$secondary_treatment, NA)


ggplot(pam_data4, aes(x = Date, y = Fv.Fm_avg, color = Intial_treatment, fill = SecondaryColor)) +
  geom_point()+
  facet_grid(Species ~ Treatmentstage)
  
  geom_errorbar(aes(ymin = chlc2_mean - chlc2_se, ymax = chlc2_mean + chlc2_se), 
                width = 0.2, position = position_dodge(width = 0.75)) +
  geom_point(position = position_dodge(width = 0.75), shape = 21, size = 3) +
  facet_grid(Species ~ TreatmentStage) +
  theme_classic()+
  scale_color_manual(
    name = "Initial Treatment",
    values = c("Control" = "lightblue", "High_Temp" = "red", "Low_pH" = "purple")) +
  scale_fill_manual(
    name = "Secondary Color",
    values = c("Control" = "blue", "High_Temp" = "darkred", "NA" = "grey")) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1),
        panel.spacing = unit(0.3, "lines"),
     panel.border = element_rect(color = "black", fill = NA)  )+
  labs(x = "Initial Treatment", y = expression("Average Chlorophyll C2 (µg/cm"^2*")"))+
ggsave("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/chlorophyll/Chlc2-3.pdf", device = "pdf", width = 8, height = 6)

pam_data3 <- pam_data3[, -c(1,2,5,7:18)]

pam_summary_avg <- pam_data3 %>%
  summarize(
    Fv.Fm_mean = mean(Fv.Fm_avg), 
    Fv.Fm_se = sqrt(sum(Fv.Fm_avg^2)/n()),
    .groups = "drop"
  )

ggplot(pam_data, aes(x = Species, y = Fv.Fm_avg)) +
  geom_boxplot(aes(colour=Treatment), width=.5, outlier.shape= NA) +
  coord_cartesian(ylim = c(-0, 0.7))+
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  #geom_violin(aes(color = Treatment)) +
  #scale_fill_manual(values=c("lightblue", "pink"))+
  scale_colour_manual(values=c("blue", "green","red","purple"))+
  theme_bw()+ theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank())
```



Plot Fv/Fm by treatment, species and colony
```{r}
ggplot(pam_data, aes(x = Species, y = Fv.Fm_avg, color = Temperature)) +
  geom_boxplot() +
  coord_cartesian(ylim = c(-0, 0.7))+
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  #geom_violin(aes(color = Treatment)) +
  #scale_fill_manual(values=c("lightblue", "pink"))+
  scale_colour_manual(values=c("blue", "black","red","purple"))+
  theme_bw() +
  facet_wrap(~ Colony)+ theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank())
  #geom_boxplot(aes(color = Treatment)) +
  #geom_violin()
#pdf("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/Pam/By_colony.pdf")
```
Do stats on Fv/Fm by treatment, species and colony 
```{r}
model <- aov(Fv.Fm_avg ~ Treatment*Species*Colony, data = pam_data)
summary(model)

# View post-hoc differences 
emm = emmeans(model, ~ Treatment*Species*Colony)
cld(emm, Letters=c(LETTERS)) #letter display
pairs(emm)

```

Plot Fv/Fm by treatment and species
```{r}
ggplot(pam_data, aes(x = Species, y = Fv.Fm_avg)) +
  geom_boxplot(aes(colour=Treatment), width=.5, outlier.shape= NA) +
  coord_cartesian(ylim = c(-0, 0.7))+
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  #geom_violin(aes(color = Treatment)) +
  #scale_fill_manual(values=c("lightblue", "pink"))+
  scale_colour_manual(values=c("blue", "green","red","purple"))+
  theme_bw()+ theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank())+
  #geom_boxplot(aes(color = Treatment)) +
  #geom_violin()
pdf("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/Pam/By_species.pdf")
```

Do stats on Fv/Fm by treatment and colony 
```{r}
model <- aov(Fv.Fm_avg ~ Treatment*Colony, data = pam_data)
summary(model)

# View post-hoc differences

emm = emmeans(model, ~ Treatment*Colony)
cld(emm, Letters=c(LETTERS)) #letter display
pairs(emm)
```

## Count observations 

Calculate the number of measurements grouped by date, treatment and colony
```{r}
summary_data <- pam_data %>%
  group_by(DateMeasured, colony, species, treatment) %>%
  summarise(observations = n())

summary_data <- pam_data %>%
  group_by(species, Treatment) %>%
  summarise(observations = n())
```

We may only be able to use the measurements with >3 data points per date, treatment and colony if that makes sense. 

Group by colony and plot so that date is the x-axis, Fv/Fm is the y-axis, colors are treatment, and shapes are colony
```{r}
test <- pam_data %>%
  group_by(Colony, DateMeasured, Treatment,species) %>%
  summarise(mean = mean(Fv.Fm),
            sd = sd(Fv.Fm)) %>%
  drop_na(sd)

# Plot
ggplot(test, aes(x=DateMeasured, y=mean, color = Treatment, shape = species)) +
  #geom_line() +
  #geom_boxplot() +
  geom_point(size = 5) + 
  scale_shape_manual(values = c(0:15))+ 
  scale_color_manual(values = c("blue", "red")) +
  theme_classic()
```
There is a violation of variance assumptions. Conduct non-parametric test 
```{r}
#kruskal.test(Fv.Fm ~ colony, data = pam_data_spat)
#kruskal.test(Fv.Fm ~ condition, data = pam_data_spat)
#kruskal.test(Fv.Fm ~ date, data = pam_data_spat)

# View post-hoc differences 
#emm = emmeans(model, ~ condition*colony*date)
#cld(emm, Letters=c(LETTERS)) #letter display
#pairs(emm)
```


