---
title: "PAM data analysis"
author: "Florence Fields"
date: "2024-06-12"
output: html_document
---

This code analyzes the pulse amplitude fluometry (PAM) data that was collected from from adult Mdec and Dlab in Bermuda 2024 as part of the ENCORE priming experiments. 

## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r}
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(car)
library(emmeans)
library(multcomp)
library(knitr)
library(kableExtra)
library(lme4)
library(GGally)
library(lmerTest)
library(cowplot)
library(Hmisc)
library(ggstatsplot)
```

## First, I am going to analyze the data from the larval samples. 

## Load data 
```{r}
pam_data <- read.csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/Pam/All_Pam_Data.csv", na.strings=c("","NA"))
head(pam_data)
tail(pam_data)
```

## Manipulate data frame 

Remove NAs from colony 
```{r}
#pam_data <- pam_data %>%
#  drop_na(Colony)
```

Set date
```{r}
pam_data$DateMeasured <- as.POSIXct(pam_data$DateMeasured, format = "%m/%d/%y")
```

rename colounm heads 
```{r}
colnames(pam_data)[colnames(pam_data) == "Fo.I."] <- "Fo.I"
colnames(pam_data)[colnames(pam_data) == "Fo.II."] <- "Fo.II"
colnames(pam_data)[colnames(pam_data) == "Fo.III."] <- "Fo.III"
colnames(pam_data)[colnames(pam_data) == "Fm.I."] <- "Fm.I"
colnames(pam_data)[colnames(pam_data) == "Fm.II."] <- "Fm.II"
colnames(pam_data)[colnames(pam_data) == "Fm.III."] <- "Fm.III"
colnames(pam_data)[colnames(pam_data) == "Y.I."] <- "Fv.Fm.I"
colnames(pam_data)[colnames(pam_data) == "Y.II."] <- "Fv.Fm.II"
colnames(pam_data)[colnames(pam_data) == "Y.III."] <- "Fv.Fm.III"
colnames(pam_data)[colnames(pam_data) == "Fo..avg."] <- "Fo.avg"
colnames(pam_data)[colnames(pam_data) == "Fm..avg."] <- "Fm.avg"
colnames(pam_data)[colnames(pam_data) == "Y.avg."] <- "Fv.Fm_avg"
```
## Make plots, identify outliers and do statistics 


Plot Fv/Fm by treatment, species and date
```{r}
ggplot(pam_data, aes(x = Treatment, y = Fv.Fm_avg, color = "Species")) +
  geom_boxplot(aes(color=Treatment), width=.5, outlier.shape= NA)+
  coord_cartesian(ylim = c(-0, 0.7))+
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  #geom_violin(aes(color = Treatment)) +
  #scale_fill_manual(values=c("lightblue", "pink"))+
  scale_colour_manual(values=c("blue", "red", "green"))+
  #theme_bw() +
  facet_wrap(~ DateMeasured)
  #geom_boxplot(aes(color = Species)) +
  #geom_violin()
```

NEED TO LOOK AT OUTLIERS 

Do stats on Fv/Fm by treatment, species and date 
```{r}
model <- aov(Fv.Fm_avg ~ Treatment*Species*DateMeasured, data = pam_data)
summary(model)

# View post-hoc differences 
emm = emmeans(model, ~ Treatment*Species*DateMeasured)
cld(emm, Letters=c(LETTERS)) #letter display
pairs(emm)

```

Plot Fv/Fm by treatment, species and colony
```{r}
ggplot(pam_data, aes(x = Species, y = Fv.Fm_avg, color = Treatment)) +
  geom_boxplot(aes(colour=Treatment,), width=.5, outlier.shape= NA) +
  coord_cartesian(ylim = c(-0, 0.7))+
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  #geom_violin(aes(color = Treatment)) +
  #scale_fill_manual(values=c("lightblue", "pink"))+
  scale_colour_manual(values=c("blue", "red", "green"))+
  theme_bw() +
  facet_wrap(~Colony)
  #geom_boxplot(aes(color = Treatment)) +
  #geom_violin()
```
Do stats on Fv/Fm by treatment, species and colony 
```{r}
model <- aov(Fv.Fm_avg ~ Treatment*Species*Colony, data = pam_data)
summary(model)

# View post-hoc differences 
emm = emmeans(model, ~ Treatment*Species*Colony)
cld(emm, Letters=c(LETTERS)) #letter display
pairs(emm)

```

Plot Fv/Fm by treatment and species
```{r}
ggplot(pam_data, aes(x = Species, y = Fv.Fm_avg)) +
  geom_boxplot(aes(colour=Treatment), width=.5, outlier.shape= NA) +
  coord_cartesian(ylim = c(-0, 0.7))+
  #geom_point(aes(fill = Treatment, group = colony), pch = 21, size = 4) +
  #geom_violin(aes(color = Treatment)) +
  #scale_fill_manual(values=c("lightblue", "pink"))+
  scale_colour_manual(values=c("blue", "red", "green"))+
  theme_bw()
  #geom_boxplot(aes(color = Treatment)) +
  #geom_violin()
```

Do stats on Fv/Fm by treatment and colony 
```{r}
model <- aov(Fv.Fm_avg ~ Treatment*Colony, data = pam_data)
summary(model)

# View post-hoc differences 
TukeyHSD(model)

emm = emmeans(model, ~ Treatment*Colony)
cld(emm, Letters=c(LETTERS)) #letter display
pairs(emm)
```

## Count observations 

Calculate the number of measurements grouped by date, treatment and colony
```{r}
summary_data <- pam_data %>%
  group_by(DateMeasured, colony, species, treatment) %>%
  summarise(observations = n())

summary_data <- pam_data %>%
  group_by(species, Treatment) %>%
  summarise(observations = n())
```

We may only be able to use the measurements with >3 data points per date, treatment and colony if that makes sense. 

Group by colony and plot so that date is the x-axis, Fv/Fm is the y-axis, colors are treatment, and shapes are colony
```{r}
test <- pam_data %>%
  group_by(Colony, DateMeasured, Treatment,species) %>%
  summarise(mean = mean(Fv.Fm),
            sd = sd(Fv.Fm)) %>%
  drop_na(sd)

# Plot
ggplot(test, aes(x=DateMeasured, y=mean, color = Treatment, shape = species)) +
  #geom_line() +
  #geom_boxplot() +
  geom_point(size = 5) + 
  scale_shape_manual(values = c(0:15))+ 
  scale_color_manual(values = c("blue", "red")) +
  theme_classic()
```
There is a violation of variance assumptions. Conduct non-parametric test 
```{r}
#kruskal.test(Fv.Fm ~ colony, data = pam_data_spat)
#kruskal.test(Fv.Fm ~ condition, data = pam_data_spat)
#kruskal.test(Fv.Fm ~ date, data = pam_data_spat)

# View post-hoc differences 
#emm = emmeans(model, ~ condition*colony*date)
#cld(emm, Letters=c(LETTERS)) #letter display
#pairs(emm)
```


