---
title: "Total Protein"
author: "EL Strand"
edited: "Florence Fields"
date: "2025-01-05"
output: html_document
---

Code modified from R. Cunning, H. Putnam, and E5 group [here](https://github.com/urol-e5/timeseries/blob/master/timepoint_1/scripts/Protein.Rmd) and [Emma Stand](https://github.com/hputnam/Acclim_Dynamics/blob/master/Scripts/Protein-Purple-Kit.Rmd)

9-10 protein plates with each one containing a standard curve. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## install packages if you dont already have them
if (!require("plyr")) install.packages("plyr")
if (!require("dplyr")) install.packages("dplyr")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("broom")) install.packages("broom")
if (!require("Rmisc")) install.packages("Rmisc")
if (!require("sjPlot")) install.packages("sjPlot")
if (!require("insight")) install.packages("insight")



# load packages
library(tidyr)
library(plyr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(tidyverse)
library(Rmisc)
library(broom)
library(lme4)
library(car)
library(sjPlot)
library(purrr)
library(readr)


```

#Read in all data files. 
```{r}

setwd("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/Protein")

prot_path <- "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/Protein/" # Path to data directory

getwd()


# Define function to read in prot data
read_prot <- function(file) {
  prot_data <- read_csv(file, skip = 23, n_max = 24)%>%
    select(-1)%>%
    magrittr::set_colnames(c("row", 1:12, "wavelength"))%>%
    fill(row)%>%
    gather("col", "absorbance", -wavelength, -row)%>%
    unite("well", c(row, col), sep = "")
}

all_prot_files <- list.files(path = prot_path, pattern = "*.csv")   # List all files in directory
prot_platemaps <- list.files(path = prot_path, pattern = "map")     # List platemap files
prot_data_files <- setdiff(all_prot_files, prot_platemaps) 

# Read in all files into tibble
df <- tibble(file = prot_data_files) %>%
  separate(file, into = c("date", "plate"), remove = FALSE) %>%
  unite(plate, date, plate) %>%
  mutate(platemap = map(plate, ~read_csv(paste0(prot_path, ., "_map.csv"))),
         prot_data = map(file, ~read_prot(paste0(prot_path, .)))) 

# Print the resulting tibble
print(df)
         
# Merge platemap and data for each plate
df <- df %>%
  mutate(merged = map2(platemap, prot_data, ~ right_join(.x, .y)))
```

Create standard curves. 
```{r}
# Create standard curve following kit instructions
standards <- tribble(
  ~std, ~BSA_ug.mL,
  "A",        2000,
  "B",        1500,
  "C",        1000,
  "D",         750,
  "E",         500,
  "F",         250,
  "G",         125,
  "H",          25,
  "I",           0
)

std_curve <- df %>%
  unnest(merged) %>%
  filter(grepl("Standard", Fragment_ID)) %>%
  select(Date_Processed, well, Fragment_ID, abs562 = `absorbance`) %>%
  dplyr::rename(std = Fragment_ID) %>%
  mutate(std = str_sub(std, 9, 9)) %>%
  dplyr::group_by(Date_Processed, std) %>%
  mutate(abs562.mean = mean(abs562)) %>% # calculates means using the above group by fxns 
  select(-well, -abs562) %>% # taking out original values to then be able to remove duplicates
  distinct() %>%
  ungroup(std) %>% # the below fxn needs to be only grouped by plate
  mutate(abs562 = abs562.mean - abs562.mean[std == "I"]) %>% # subtract blank absorbance value from all (adjusted value)
  left_join(standards)
```

Fit a linear model for the standards. 
```{r}
mod <- lm(formula = BSA_ug.mL ~ abs562, data = std_curve)
fitted <- mod %>% broom::augment()
std_curve <- left_join(std_curve, fitted)

# Plot standard curve
std_curve_plot <- 
  ggplot(aes(x = abs562, y = BSA_ug.mL, group=Date_Processed, color=Date_Processed), data = std_curve) +
  geom_point()

std_curve_plot + 
  geom_line(data = std_curve, aes(x = abs562, y = .fitted)) +
  labs(title = "Standard curve") + 
  theme_classic()
```

Calculate protein concentration for all samples using standard curve
```{r}
prot <- df %>%
  unnest(merged) %>%
  filter(!grepl("Standard", Fragment_ID)) %>%                     # Get just samples (not standards)
  select(Date_Processed, well, Fragment_ID, abs562 = `absorbance`) %>%        # Select only needed columns
  filter(!is.na(Fragment_ID)) %>%                                 # Filter out empty wells
  mutate(prot_ug.mL = map_dbl(abs562, ~ predict(mod, newdata = data.frame(abs562 = .)))) # Use standard curve to convert absorbance to protein 
  
std_curve_plot + 
  geom_point(data = prot, aes(x = abs562, y = prot_ug.mL), pch = "X", cex = 5, alpha = 0.3) +
  labs(title = "All samples projected on standard curve")

prot <- prot %>% ## I keep this after the std curve plot so I can plot the abs562 above and take that column out below 
  group_by(Date_Processed, Fragment_ID) %>%
  mutate(prot_ug.mL = mean(prot_ug.mL)) %>% select(-well, -abs562) %>% distinct()

## visual check to see if any protein samples are above the standard curve 
## if so, I would have to dilute them and redo the sample 
```

Normalize to surface area
```{r}
# Surface area data
sa <- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/Surface_Area/Surface_Area.csv") %>% dplyr::rename(Fragment_ID = Colony_id) %>% select(Fragment_ID, surface.area.cm2)
sa$Fragment_ID <- as.character(sa$Fragment_ID)

# Tissue homogenate volume data
homog_vols <- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/Airbrushing_data.csv") %>% select(Species, Fragment_ID, Homogenate.Vol,Tank
)
homog_vols$Fragment_ID <- as.character(homog_vols$Fragment_ID)

data <- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/Coral_Fragment_Metadata1.csv") %>% select(Sample.ID, Secondary.Treatment) %>% dplyr::rename(Fragment_ID = Sample.ID)

# Join homogenate volumes and surface area with sample metadata
metadata <- full_join(data, homog_vols)%>% full_join(sa)%>% na.omit() 


# Join prot data with metadata
prot <- left_join(prot, metadata) %>%
  mutate(prot_ug = prot_ug.mL * Homogenate.Vol,
         prot_ug.cm2 = prot_ug / surface.area.cm2,
         prot_mg.cm2 = prot_ug.cm2 / 1000)

## data vis and outlier check 
boxplot(prot$prot_ug)
hist(prot$prot_ug) 

boxplot.stats(prot$prot_ug)$out


prot %>% ggplot(aes(x=as.character(Date_Processed), y=prot_ug)) + geom_boxplot() #looking for any plate that is an outlier 


prot %>% filter(!is.na(Species)) %>% select(-Date_Processed)%>%
  write_csv(., path = "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/Protein/Soluble_Protein.csv")

```
Including Treatments and Calculating mean
```{r}
prot <- prot %>%
  mutate(Priming_Treatment = NA_character_) %>%  # Create empty column if missing
  mutate(Priming_Treatment = case_when(
    Tank %in% c(6, 7) ~ "High Heat",
    Tank %in% c(2, 3) ~ "Low pH",
    Tank %in% c(4, 5) ~ "Control",
    Tank %in% c("acclimation") ~ "Acclimation",
    Tank %in% c("start of acclimation")~ "Start of Acclimation",
    TRUE ~ Priming_Treatment  # Keep other values unchanged
  ))

prot <- prot %>%
  mutate(Sample.Date = as.Date(as.character(Sample.Date), format = "%Y%m%d"))

prot <- prot %>%
  dplyr::mutate(Treatmentstage = case_when(
    Sample.Date == "2024-08-06" ~ "Recovery",
    Sample.Date == "2024-08-15" ~ "Stress",
    Sample.Date == "2024-06-20" ~ "Acclimation",
    Sample.Date == "2024-07-09" ~ "Acclimation",
    Sample.Date == "2024-07-23" ~ "Priming",
    TRUE ~ as.character(Sample.Date)
  ))


```

Calculating the mean and SE
```{r}
df_summary <- prot %>%
  group_by(Species, Treatmentstage,Priming_Treatment,Secondary.Treatment) %>%
  summarize(
    prot_mean = mean(prot_mg.cm2),
    prot_se = sd(prot_mg.cm2) / sqrt(n())
  )

#Removing Acclimation rows
df_summary <- df_summary[-c(1, 2, 3, 17), ]

df_summary <- df_summary %>%
  mutate(Priming_Treatment = case_when(
    Priming_Treatment == "High Heat" ~ "High Temp",
    TRUE ~ Priming_Treatment  # Keeps other values unchanged
  ))

df_summary <- df_summary %>%
  mutate(Secondary.Treatment = case_when(
    Secondary.Treatment == "High_Temp" ~ "High Temp",
    TRUE ~ Secondary.Treatment  # Keeps other values unchanged
  ))


```

Plot te graph
```{r}
ggplot(df_summary, aes(x = Priming_Treatment, y = prot_mean, 
                         color = ifelse(Treatmentstage == "Stress", Secondary.Treatment, "black"), 
                         fill = ifelse(Treatmentstage == "Stress", Secondary.Treatment, "black"),
                         group = Priming_Treatment)) + 
  geom_errorbar(aes(ymin = prot_mean - prot_se, ymax = prot_mean + prot_se), 
                width = 0.3, position = position_dodge(width = 0.75), linewidth = 0.5) + 
  geom_point(aes(shape = Priming_Treatment), position = position_dodge(width = 0.75), size = 3) + 
  facet_grid(Species ~ Treatmentstage, scales = "free_x") + 
  theme_classic() + 
  scale_color_manual(name = "Secondary Treatment", values = c("Control" = "blue", "High Temp" = "darkred", "Low pH" = "purple", "NA" = "black")) + 
  scale_fill_manual(name = "Secondary Treatment",
                    values = c("Control" = "blue", "High Temp" = "darkred", "Low pH" = "purple", "NA" = "black")) + 
  scale_shape_manual(name = "Primary Treatment", values = c("Control" = 16, "High Temp" = 17, "Low pH" = 18)) + 
  theme(axis.text.x = element_text(angle = 35, hjust = 1), 
      panel.spacing = unit(0.3, "lines"), 
      panel.border = element_rect(color = "black", fill = NA)) +
labs(x = "Priming Treatment", y = expression("Average Protein (µg/cm"^2*")"))
ggsave("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/Protein/protein_point.pdf", device = "pdf", width = 8, height = 6)

```

Statistics
```{r}
colnames(prot)[15] <- "TreatmentStage"
DL <- prot %>%
  filter(!grepl("Madracis decactis", Species))

#DL STATS

model <- aov(prot_mg.cm2 ~ Priming_Treatment*Secondary_Treatment*TreatmentStage, data = DL)
summary(model)

```
One way - Dlab

```{r}

MD <- prot %>%
  filter(!grepl("Diploria labyrinthiformis", Species))


model <- aov(prot_mg.cm2 ~ Priming_Treatment*Secondary_Treatment*TreatmentStage, data= MD)
summary(model)

```
Mdec

