---
title: "Daily measurement analysis of 2024 Priming experiment ENCORE"
author: "Florence Fields"
date: "2024-06-12"
output: html_document
---

This script reads and plots environmental data from daily measurements. This script was created by Jill Ashey and altered to suit this data and experiment purposes.

## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

Load packages

```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(purrr)
library(lubridate)
library(ggplot2)
library(seacarb)
library(broom)
library(cowplot)
```

Read in daily measurements file.

```{r}
daily<-read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/daily_measurements.csv")
daily$date<-as.Date(daily$date, format="%m/%d/%y")
daily$tris.date<-as.character(daily$tris.date)

# remove values with NA for temp, salinity and pH
daily <- daily %>%
  drop_na(temp.C)
```

## Combine tris cal datasets

```{r setup, echo=FALSE}
tris1<- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/Tris/20240619.csv")
tris2<- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/Tris/20240626.csv")
tris3<- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/Tris/20240701.csv")
tris4<- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/Tris/20240704.csv")
tris5<- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/Tris/20240707.csv")
tris6<- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/Tris/20240710.csv")
tris7<- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/Tris/20240715.csv")
tris8<- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/Tris/20240718.csv")
tris9<- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/Tris/20240724.csv")
tris10<- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/Tris/20240807.csv")

tris_6<-tris6[-c(1:3), ]
tris_list<- list(tris1,tris2,tris3,tris4,tris5,tris_6,tris7,tris8,tris9,tris10)

tris_merge <-Reduce(function(x, y) merge(x, y, all=TRUE), tris_list)

write.csv(tris_merge, "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/environmental/TrisCal/Tris_Calibration.csv")
```

## Calculate total pH

Calculate the calibration curve from the Tris calibration and calculate pH on the total scale from pH.mV.

```{r}
pHcalib<-read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/environmental/TrisCal/Tris_Calibration.csv")
pHcalib$Tris.date<-as.character(pHcalib$Tris.date)

daily <- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/daily_measurements.csv")

#Changing the name of coloumn 11 in the daily dataset from tris.date to Tris.date
colnames(daily)[colnames(daily) == 'tris.date'] <- 'Tris.date'


pHSlope <- pHcalib %>%
  nest_by(Tris.date) %>%
  mutate(fitpH = list(lm(mVTris ~ Ttris, data = pHcalib))) %>% # linear regression of mV and temp of the tris
  reframe(broom::tidy(fitpH)) %>% # make the output tidy
  dplyr::select(Tris.date, term, estimate) %>%
  pivot_wider(names_from = term, values_from = estimate) %>% # put slope and intercept in their own column
  mutate(Tris.date = as.character(Tris.date)) %>% # convert Tris.date to character
  left_join(daily %>% mutate(Tris.date = as.character(Tris.date)), by = "Tris.date") %>% # join with the pH sample data
  mutate(mVTris = temp.C * Ttris + `(Intercept)`) %>% # calculate the mV of the tris at temperature in which the pH of the tanks were measured
  mutate(pH.total = seacarb::pH(Ex = pH.mV, Etris = mVTris, S = sal.psu, T = temp.C)) # calculate pH of the tanks using the pH seacarb function

```

pH is now calculated as Total pH in the "pH" column. Now select the desired columns to analyze further.

```{r}
daily_calc<-pHSlope%>%
  dplyr::select(date, time, tank, treatment, temp.C, sal.psu, par.avg, flow.mL10s, pH.nbs, pH.total)
```

## Calculate flow and par

Calculate flow to total mL per minute rather than 5 sec.

```{r}
daily_calc<-daily_calc%>%
   mutate(flow.L.min=(flow.mL10s*12)/1000)%>%
   dplyr::select(!flow.mL10s)
```

No need to apply the immersion correction factor. The MQ-510 Underwater Quantum Meters already apply the immersion effect correction factor to the meter readings through the meter firmware. [See Documentation Here](https://www.apogeeinstruments.com/underwater-par-measurements/)

## Remove nbs values and change to long format

Change data format to long format

```{r}
daily_calc <- daily_calc %>% filter(pH.total >= 6)
daily_calc <- daily_calc %>% filter(sal.psu >=35)
daily_calc <- daily_calc %>% mutate(treatment = ifelse(treatment == "contrrol", "control", treatment))
daily_calc <- daily_calc[, -which(names(daily_calc) == "pH.nbs")]
daily_calc.long <-daily_calc %>% pivot_longer(cols=temp.C:flow.L.min,
  names_to = "metric",
  values_to = "value")

```

Filter by relevant dates if needed

# Convert continuous variable to categorical

#data$category <- cut(data$value, breaks = c(0, 3, 6, 9), labels = c("Low", "Medium", "High"))

## 

```         
```

## Plot metrics of interest

Plot by date colored by treatment and tank. (may change script in the fields according to my needs)

```{r}
daily_calc.long <- daily_calc.long %>% arrange(treatment, date)
## Treatment
daily_treatment<-daily_calc.long %>%
  filter(!is.na(value)) %>%
  ggplot(aes(x=date, y=value, colour=treatment))+
  geom_point(size=2)+
  xlab("Date")+
  facet_wrap(~metric,, scales= "free_y", nrow=5)+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  theme_bw(); daily_treatment
  
ggsave(filename="C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/environmental/treatment_daily_measurements.pdf", plot=daily_treatment, dpi=300, width=6, height=8, units="in")


daily_treatment<-daily_calc.long %>%
  filter(!is.na(value)) %>%
  ggplot(aes(x=date, y=value, colour=treatment))+
  geom_line(size=.5)+
  xlab("Date")+
  theme(axis.text.x = element_text(angle=50, hjust=1))+
  facet_wrap(metric ~ ., scales = "free")+
  theme_bw(); daily_treatment
  
ggsave(filename="C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/environmental/treatment_daily_measurements_line_graph.pdf", plot=daily_treatment, dpi=300, width=6, height=8, units="in")


## Tank type
daily_tank2 <- daily_calc.long %>%
  filter(!is.na(value)) %>%  
  ggplot(aes(x = date, y = value, colour = tank)) +
  geom_point(size = 2) +
   scale_color_gradientn(colours = c("blue", "darkgreen", "yellow", "red"))+
  xlab("Date") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 65, hjust = 1)) +
  facet_grid(metric ~ ., scales = "free")


ggsave(filename= "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/environmental/tank_daily_measurements2_.pdf", plot=daily_tank2, dpi=300, width=6, height=8, units="in")

```

```{r}
library(dplyr)
library(ggplot2)

# Ensure the date column is in Date format
daily_calc.long$date <- as.Date(daily_calc.long$date, format = "%Y-%m-%d")

# Check for missing values
sum(is.na(daily_calc.long$date))
sum(is.na(daily_calc.long$value))
sum(is.na(daily_calc.long$value))

# Ensure the data is sorted by date within each treatment
daily_calc.long <- daily_calc.long %>%
  arrange(treatment, date)

# Plot the data
daily_calc.long %>% filter(!is.na(value)) %>% ggplot(aes(x = date, y = value, colour = treatment)) + geom_point(size = 2) + xlab("Date") + facet_wrap(metric ~ ., scales = "free_y", nrow = 5) + theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + theme_bw()

# Display the plot
print(daily_treatment)

```

## Test for differences

Use the daily_calc df created earlier.

Test for differences between treatments.

```{r}
daily_calc%>%
  aov(flow.L.min~treatment, data=.)%>%
  summary()
 
daily_calc%>%
  aov(par.avg~treatment, data=.)%>%
  summary()

daily_calc%>%
  aov(pH.total~treatment, data=.)%>%
  summary()

daily_calc%>%
  aov(pH.nbs~treatment, data=.)%>%
  summary()

daily_calc%>%
  aov(sal.psu~treatment, data=.)%>%
  summary()

daily_calc%>%
  aov(temp.C~treatment, data=.)%>%
  summary()
```

Test for differences between tanks and treatment

```{r}
daily_calc%>%
  aov(flow.L.min~tank*treatment, data=.)
 
daily_calc%>%
  aov(par~tank*treatment, data=.)%>%
  summary()

daily_calc%>%
  aov(pH.total~tank*treatment, data=.)%>%
  summary()

daily_calc%>%
  aov(pH.nbs~tank*treatment, data=.)%>%
  summary()

daily_calc%>%
  aov(sal.psu~tank*treatment, data=.)%>%
  summary()

daily_calc%>%
  aov(temp.C~tank*treatment, data=.)%>%
  summary()
```

## Summarize daily measurements

Calculate descriptive statistics

```{r}
summary<-daily_calc%>%
  group_by(tank, treatment)%>%
  select(!date)%>%
  select(!time)%>%
  summarise(across(everything(), list(mean = mean, sd = sd), na.rm = TRUE)); summary

write_csv(summary, "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/environmental/daily_measurements_summary.csv")
```
