---
title: "Daily measurement analysis of 2024 Priming experiment ENCORE"
author: "Florence Fields"
date: "2024-06-12"
output: html_document
---

This script reads and plots environmental data from daily measurements. This script was created by Jill Ashey and altered to suit this data and experiment purposes.

## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

Load packages

```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(purrr)
library(lubridate)
library(ggplot2)
library(seacarb)
library(broom)
library(cowplot)
```

Read in daily measurements file.

```{r}
daily<-read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/daily_measurements.csv")
daily$date<-as.Date(daily$date, format="%m/%d/%y")
daily$tris.date<-as.character(daily$tris.date)

# remove values with NA for temp, salinity and pH
daily <- daily %>%
  drop_na(temp.C)
```

## Combine tris cal datasets

```{r setup, echo=FALSE}
tris1<- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/Tris/20240619.csv")
tris2<- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/Tris/20240626.csv")
tris3<- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/Tris/20240701.csv")
tris4<- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/Tris/20240704.csv")
tris5<- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/Tris/20240707.csv")
tris6<- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/Tris/20240710.csv")
tris7<- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/Tris/20240715.csv")
tris8<- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/Tris/20240718.csv")
tris9<- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/Tris/20240724.csv")
tris10<- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/Tris/20240807.csv")

tris_6<-tris6[-c(1:3), ]
tris_list<- list(tris1,tris2,tris3,tris4,tris5,tris_6,tris7,tris8,tris9,tris10)

tris_merge <-Reduce(function(x, y) merge(x, y, all=TRUE), tris_list)

write.csv(tris_merge, "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/environmental/TrisCal/Tris_Calibration.csv")
```

## Calculate total pH

Calculate the calibration curve from the Tris calibration and calculate pH on the total scale from pH.mV.

```{r}
pHcalib<-read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/environmental/TrisCal/Tris_Calibration.csv")
pHcalib$Tris.date<-as.character(pHcalib$Tris.date)

daily <- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/daily_measurements.csv")

#Changing the name of coloumn 11 in the daily dataset from tris.date to Tris.date
colnames(daily)[colnames(daily) == 'tris.date'] <- 'Tris.date'


pHSlope <- pHcalib %>%
  nest_by(Tris.date) %>%
  mutate(fitpH = list(lm(mVTris ~ Ttris, data = pHcalib))) %>% # linear regression of mV and temp of the tris
  reframe(broom::tidy(fitpH)) %>% # make the output tidy
  dplyr::select(Tris.date, term, estimate) %>%
  pivot_wider(names_from = term, values_from = estimate) %>% # put slope and intercept in their own column
  mutate(Tris.date = as.character(Tris.date)) %>% # convert Tris.date to character
  left_join(daily %>% mutate(Tris.date = as.character(Tris.date)), by = "Tris.date") %>% # join with the pH sample data
  mutate(mVTris = temp.C * Ttris + `(Intercept)`) %>% # calculate the mV of the tris at temperature in which the pH of the tanks were measured
  mutate(pH.total = seacarb::pH(Ex = pH.mV, Etris = mVTris, S = sal.psu, T = temp.C)) # calculate pH of the tanks using the pH seacarb function

```

pH is now calculated as Total pH in the "pH" column. Now select the desired columns to analyze further.

```{r}
daily_calc<-pHSlope%>%
  dplyr::select(date, time, tank, treatment, temp.C, sal.psu, par.avg, flow.mL10s, pH.nbs, pH.total)
```

## Calculate flow and par

Calculate flow to total mL per minute rather than 5 sec.

```{r}
daily_calc<-daily_calc%>%
   mutate(flow.L.min=(flow.mL10s*12)/1000)%>%
   dplyr::select(!flow.mL10s)
```

No need to apply the immersion correction factor. The MQ-510 Underwater Quantum Meters already apply the immersion effect correction factor to the meter readings through the meter firmware. [See Documentation Here](https://www.apogeeinstruments.com/underwater-par-measurements/)

## Remove nbs values and change to long format

Change data format to long format

```{r}
daily_calc <- daily_calc %>% filter(pH.total >= 6)
daily_calc <- daily_calc %>% filter(sal.psu >=35)
daily_calc <- daily_calc %>% mutate(treatment = ifelse(treatment == "contrrol", "control", treatment))
daily_calc <- daily_calc[, -which(names(daily_calc) == "pH.nbs")]
daily_calc.long <-daily_calc %>% pivot_longer(cols=temp.C:flow.L.min,
  names_to = "metric",
  values_to = "value")

```

Filter by relevant dates if needed

# Convert continuous variable to categorical

#data$category <- cut(data$value, breaks = c(0, 3, 6, 9), labels = c("Low", "Medium", "High"))

## 

```         
```

## Plot metrics of interest

Plot by date colored by treatment and tank. (may change script in the fields according to my needs)

```{r}
daily_calc.long <- daily_calc.long %>% arrange(treatment, date)
## Treatment
daily_treatment<-daily_calc.long %>%
  filter(!is.na(value)) %>%
  ggplot(aes(x=date, y=value, colour=treatment))+
  geom_point(size=2)+
  xlab("Date")+
  facet_wrap(~metric,, scales= "free_y", nrow=5)+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  theme_bw(); daily_treatment
  
ggsave(filename="C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/environmental/treatment_daily_measurements.pdf", plot=daily_treatment, dpi=300, width=6, height=8, units="in")


#renaming contorrl to control
daily_calc.long <- daily_calc.long %>%
  mutate(treatment = case_when(
    treatment == "contrrol" ~ "control",
    TRUE ~ treatment  # Keep the original value if no match
  ))

daily_treatment<-daily_calc.long %>%
  filter(!is.na(value)) %>%
  ggplot(aes(x=date, y=value, colour=treatment))+
  geom_line(size=.5)+
  xlab("Date")+
  theme(axis.text.x = element_text(angle=50, hjust=1))+
  facet_wrap(metric ~ ., scales = "free")+
  theme_bw(); daily_treatment


#Plotting a line graph of temp 

daily_calc.long2 <- daily_calc.long %>%
  filter(metric != "flow.L.min",
         metric != "sal.psu",
         metric != "par.avg",
         metric != "pH.total")
# Convert the date strings to Date type
daily_calc.long2$date <- as.Date(daily_calc.long2$date, format = "%m/%d/%Y")  # Adjust the format string as necessary


#Combining tank and treatment variables to create new column  
daily_calc.long2 <- daily_calc.long2 %>%
  mutate(tank_intial_treatment = paste(tank, treatment))

daily_calc.long2 <- daily_calc.long2 %>%
  mutate(tank_intial_treatment = case_when(
    tank_intial_treatment == "6 low_ph" ~ "6 heated",
    TRUE ~ tank_intial_treatment
  ))

daily_calc.long2 <- daily_calc.long2 %>%
  mutate(tank_intial_treatment = case_when(
    tank_intial_treatment == "7 low_ph" ~ "7 heated",
    TRUE ~ tank_intial_treatment
  ))



ggplot(daily_calc.long2, aes(date, value, color = tank_intial_treatment))+
  geom_line()+
  scale_fill_manual(name = "Tank Treatment", 
                    values = c("lightblue", "blue", "darkblue", "red", "darkred","magenta",  "purple"))+
  scale_color_manual(label = c("Acclimation","Tank 2 = Low pH", "Tank 3 = Low pH","Tank 4 = Control", "Tank 5 = Control", "Tank 6 = High Heat", "Tank 7 = High Heat"), 
                     name = "Initial Tank Treatment", values = c("lightblue", "blue", "darkblue", "red", "darkred","magenta",  "purple"))+
    labs(x = "Date", y = "Temperature °C")+theme_classic()
ggsave(filename = "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/environmental/temp_daily_line_graph.pdf", device = "pdf", width = 10, height = 10)


## Tank type
daily_tank2 <- daily_calc.long %>%
  filter(!is.na(value)) %>%  
  ggplot(aes(x = date, y = value, colour = tank)) +
  geom_point(size = 2) +
   scale_color_gradientn(colours = c("blue", "darkgreen", "yellow", "red"))+
  xlab("Date") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 65, hjust = 1)) +
  facet_grid(metric ~ ., scales = "free")

ggsave(filename= "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/environmental/tank_daily_measurements2_.pdf", plot=daily_tank2, dpi=300, width=6, height=8, units="in")

```

# Combining Apex and filling in missing data with daily measurements 

pH Measurements

```{r}
#pH measurements 


daily_calc.long3 <- daily_calc.long %>%
  filter(metric != "flow.L.min",
         metric != "sal.psu",
         metric != "par.avg",
         metric != "temp.C")
# Convert the date strings to Date type
daily_calc.long3$date <- as.Date(daily_calc.long3$date, format = "%m/%d/%Y")  # Adjust the format string as necessary

#import ph apex data
apex <- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/environmental/Apex/Merged_Apex_Data_Output.csv")

#removing temp colunms
apexpH <- apex %>%
  select(-c(1,3,5,7,9,11,13,15))

#seperating my date,time colunm to a date and time colunm
apexpH <- apexpH %>%
  separate(Date.Time, into = c("date", "time"), sep = " ")


#convert datset to long format
apexpH_long <- apexpH %>%
  pivot_longer(cols = starts_with("pH_T"),
               names_to = "tank_treatment",
               values_to = "value") %>%
  separate(tank_treatment, into = c("pH", "tank_treatment"), sep = "_") %>%
  select(-pH)

apexpH_long <- apexpH_long %>%
  mutate(tank_intial_treatment = case_when(
    tank_treatment == "T2" ~ "2 low_ph",
    tank_treatment == "T3" ~ "3 low_ph",
    tank_treatment == "T4" ~ "4 control",
    tank_treatment == "T5" ~ "5 control",
    tank_treatment == "T6" ~ "6 heated",
    tank_treatment == "T7" ~ "7 heated",
    TRUE ~ tank_treatment
  ))

apexpH_long <- apexpH_long%>%
  select(date, time, value, tank_intial_treatment ) %>%
  mutate(date = format(as.Date(date, format = "%m/%d/%Y"), "%Y-%m-%d"))

average_values <- apexpH_long%>%
  group_by(date, tank_intial_treatment)%>%
  summarise(avg_value = mean(value, na.rm = TRUE))

#removing unwanted rows
average_values <- average_values[-c(175:180), ]

#remove pH data from daily_calc.long3 that will be replaced
daily_calc.long4 <- daily_calc.long3[-c(10:35,38:41, 50:55, 63:95, 98:101, 110:115, 120:155, 158:161, 170:175, 184:193), ]


daily_calc.long4 <- daily_calc.long4 %>%
  mutate(treatment = case_when(
    tank == "6" ~ "High Temp",
    tank == "7" ~ "High Temp",
    tank == "2" ~ "Low pH",
    tank == "3" ~ "Low pH",
    tank == "5" ~ "Control",
    tank == "4" ~ "Control",
    tank == "1"~ "Acclimation",
    TRUE ~ as.character(tank)))


average_values <- average_values %>%
  rename(value = avg_value)

average_values <- average_values %>%
  mutate(treatment = case_when(
    tank_intial_treatment == "6 heated" ~ "High Temp",
    tank_intial_treatment == "7 heated" ~ "High Temp",
    tank_intial_treatment == "2 low_ph" ~ "Low pH",
    tank_intial_treatment == "3 low_ph" ~ "Low pH",
    tank_intial_treatment == "5 control" ~ "Control",
    tank_intial_treatment == "4 control" ~ "Control",
    TRUE ~ as.character(tank_intial_treatment)))

average_values <- average_values %>%
  mutate(tank_intial_treatment = case_when(
    tank_intial_treatment == "6 heated" ~ "6",
    tank_intial_treatment == "7 heated" ~ "7",
    tank_intial_treatment == "2 low_ph" ~ "2",
    tank_intial_treatment == "3 low_ph" ~ "3",
    tank_intial_treatment == "5 control" ~ "5",
    tank_intial_treatment == "4 control" ~ "4",
    TRUE ~ as.character(tank_intial_treatment)))

average_values <- average_values %>%
  rename(tank = tank_intial_treatment)

#average_values <- average_values[-c(170), ]

daily_calc.long4<- daily_calc.long4%>%
  select(date,tank,treatment,value)


#combinding apex and daily pH data
phuse<- rbind(daily_calc.long4,average_values)

#Adding Acclimation, Priming, recovery and Stress to the dataset
phuse<- phuse %>%
  mutate(Timepoint = case_when(
    date >= as.Date('2024-06-26') & date <= as.Date('2024-07-08') ~ 'acclimation',
    date >= as.Date('2024-07-09') & date <= as.Date('2024-07-23') ~ 'priming',
    date >= as.Date('2024-07-24') & date <= as.Date('2024-08-06') ~ 'recovery',
    date >= as.Date('2024-08-07') & date <= as.Date('2024-08-15') ~ 'stress',
    TRUE ~ 'unknown'
  ))

#Plot showing the average total pH with only 4 treatments
ggplot(phuse, aes(date, value, color = treatment))+
  geom_line(size = 1)+
  scale_color_manual(name = "Treatment", 
                    values = c("lightblue", "darkblue", "red", "purple"))+
  theme_classic()+
    labs(x = "Date", y = "Total average pH")+
ggsave(filename = "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/environmental/average_pH_treatments.pdf", device = "pdf", width = 10, height = 10)

#Plot showing the average total pH by tank
ggplot(phuse, aes(date, value, color = tank, fill = "Timpoint"))+
  geom_line(size = 1)+
  scale_color_manual(name = "Tank = Priming Treatment", 
                    values = c("lightblue", "purple", "magenta", "blue", "darkblue", "red", "darkred"),
                    labels = c("Acclimation", "2 = Low pH", "3 = Low pH", "4 = Control", "5 = Control", "6 = High Temp", "7 = High Temp")
  )+theme_classic()+
    labs(x = "Date", y = "Total Average pH")
ggsave(filename = "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/environmental/average_pH_treatments2.pdf", device = "pdf", width = 10, height = 10)
```

Temperature Measurements

```{r}

daily_calc.long2 

# Convert the date strings to Date type
daily_calc.long2$date <- as.Date(daily_calc.long2$date, format = "%m/%d/%Y")  # 

#removing temp colunms
apexTemp <- apex %>%
  select(-c(1,3,4,6,8,10,12,14,16))

#seperating my date,time colunm to a date and time colunm
apexTemp <- apexTemp %>%
  separate(Date.Time, into = c("date", "time"), sep = " ")


#convert datset to long format
apexTemp_long <- apexTemp %>%
  pivot_longer(cols = starts_with("TMP_T"),
               names_to = "tank_treatment",
               values_to = "value") %>%
  separate(tank_treatment, into = c("temp", "tank_treatment"), sep = "_")%>%
  select(-temp)

apexTemp_long <- apexTemp_long %>%
  mutate(tank_intial_treatment = case_when(
    tank_treatment == "T2" ~ "2 low_ph",
    tank_treatment == "T3" ~ "3 low_ph",
    tank_treatment == "T4" ~ "4 control",
    tank_treatment == "T5" ~ "5 control",
    tank_treatment == "T6" ~ "6 heated",
    tank_treatment == "T7" ~ "7 heated",
    TRUE ~ tank_treatment
  ))

apexTemp_long <- apexTemp_long%>%
  select(date, time, value, tank_intial_treatment ) %>%
  mutate(date = format(as.Date(date, format = "%m/%d/%Y"), "%Y-%m-%d"))

averageT_values <- apexTemp_long%>%
  group_by(date, tank_intial_treatment)%>%
  summarise(avg_value = mean(value, na.rm = TRUE))

#removing unwanted rows
averageT_values <- averageT_values[-c(176,177,179,180), ]

#remove pH data from daily_calc.long3 that will be replaced
daily_calc.long5 <- daily_calc.long2[-c(61,180), ]


daily_calc.long5 <- daily_calc.long5 %>%
  mutate(treatment = case_when(
    tank == "6" ~ "High Temp",
    tank == "7" ~ "High Temp",
    tank == "2" ~ "Low pH",
    tank == "3" ~ "Low pH",
    tank == "5" ~ "Control",
    tank == "4" ~ "Control",
    tank == "1"~ "Acclimation",
    TRUE ~ as.character(tank)))


averageT_values <- averageT_values %>%
  rename(value = avg_value)

averageT_values <- averageT_values %>%
  mutate(treatment = case_when(
    tank_intial_treatment == "6 heated" ~ "High Temp",
    tank_intial_treatment == "7 heated" ~ "High Temp",
    tank_intial_treatment == "2 low_ph" ~ "Low pH",
    tank_intial_treatment == "3 low_ph" ~ "Low pH",
    tank_intial_treatment == "5 control" ~ "Control",
    tank_intial_treatment == "4 control" ~ "Control",
    TRUE ~ as.character(tank_intial_treatment)))

averageT_values <- averageT_values %>%
  mutate(tank_intial_treatment = case_when(
    tank_intial_treatment == "6 heated" ~ "6",
    tank_intial_treatment == "7 heated" ~ "7",
    tank_intial_treatment == "2 low_ph" ~ "2",
    tank_intial_treatment == "3 low_ph" ~ "3",
    tank_intial_treatment == "5 control" ~ "5",
    tank_intial_treatment == "4 control" ~ "4",
    TRUE ~ as.character(tank_intial_treatment)))

averageT_values <- averageT_values %>%
  rename(tank = tank_intial_treatment)

#average_values <- average_values[-c(170), ]

daily_calc.long5<- daily_calc.long5  %>%
  select(date,tank,treatment,value)


#combinding apex and daily pH data
Tempuse<- rbind(daily_calc.long5,averageT_values)

#Adding Acclimation, Priming, recovery and Stress to the dataset
Tempuse<- Tempuse %>%
  mutate(Timepoint = case_when(
    date >= as.Date('2024-06-26') & date <= as.Date('2024-07-08') ~ 'acclimation',
    date >= as.Date('2024-07-09') & date <= as.Date('2024-07-24') ~ 'priming',
    date >= as.Date('2024-07-25') & date <= as.Date('2024-08-06') ~ 'recovery',
    date >= as.Date('2024-08-07') & date <= as.Date('2024-08-15') ~ 'stress',
    TRUE ~ 'unknown'
  ))

#Plot showing the average total pH with only 4 treatments
ggplot(Tempuse, aes(date, value, color = treatment))+
  geom_line(size = 1)+
  scale_color_manual(name = "Treatment", 
                    values = c("lightblue", "darkblue", "red", "purple"))+
  theme_classic()+
    labs(x = "Date", y = "Average Temperature °C")
ggsave(filename = "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/environmental/average_temp_treatments.pdf", device = "pdf", width = 10, height = 10)


#Plot showing the average total pH with only 4 treatments
ggplot(Tempuse, aes(date, value, color = tank, fill= Timepoint))+
  geom_line(size = 1)+
  scale_y_continuous(limits = c(25, 33), breaks = seq(25, 33, by = 2))+
  scale_color_manual(name = "Tank = Priming Treatment", 
                    values = c("lightblue", "purple", "magenta", "blue", "darkblue", "red", "darkred"),
                    labels = c("Acclimation", "2 = Low pH", "3 = Low pH", "4 = Control", "5 = Control", "6 = High Temp", "7 = High Temp")
  )+
  theme_classic()+
    labs(x = "Date", y = "Average Temperature °C")
ggsave(filename = "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/environmental/average_temp_treatments2.pdf", device = "pdf", width = 10, height = 10)

```

```{r}
library(dplyr)
library(ggplot2)

# Ensure the date column is in Date format
daily_calc.long$date <- as.Date(daily_calc.long$date, format = "%Y-%m-%d")

# Check for missing values
sum(is.na(daily_calc.long$date))
sum(is.na(daily_calc.long$value))
sum(is.na(daily_calc.long$value))

# Ensure the data is sorted by date within each treatment
daily_calc.long <- daily_calc.long %>%
  arrange(treatment, date)

# Plot the data
daily_calc.long %>% filter(!is.na(value)) %>% ggplot(aes(x = date, y = value, colour = treatment)) + geom_point(size = 2) + xlab("Date") + facet_wrap(metric ~ ., scales = "free_y", nrow = 5) + theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + theme_bw()

# Display the plot
print(daily_treatment)

```

## Combining daily pH and temp tank measurements

#Combining daily pH and temp

```{r}

#Combining tank temp values based on treatment
df_summary_pH_tank_avg <- daily_calc.long2 %>%
  mutate(tank_treatment = case_when(
    tank_intial_treatment %in% c("4 control", "5 control") ~ "Tank_4_5",
    tank_intial_treatment %in% c("6 heated", "7 heated") ~ "Tank_6_7",
    tank_intial_treatment %in% c("2 low_ph", "3 low_ph") ~ "Tank_2_3",
    TRUE ~ tank_intial_treatment  # Keep other tank groups unchanged
  )) %>%
  group_by(date, tank_treatment) %>%
  summarize(
    ph_mean = mean(value, na.rm = TRUE),
    ph_se = sd(value, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )


#Combining tank pH values based on treatment and adding a treatment coloumn

df_summary_pH_tank_avg <- phuse %>%
  mutate(tank_treatment = case_when((
    date >= as.Date("2024-07-10") & date <= as.Date("2024-08-06")) &
      tank_intial_treatment %in% c("4 control", "5 control") ~ "Tank_4_5",
    (date >= as.Date("2024-07-10") & date <= as.Date("2024-08-06")) &
      tank_intial_treatment %in% c("6 heated", "7 heated") ~ "Tank_6_7",
    (date >= as.Date("2024-07-10") & date <= as.Date("2024-08-06")) &
      tank_intial_treatment %in% c("2 low_ph", "3 low_ph") ~ "Tank_2_3",
    TRUE ~ tank_intial_treatment
))%>%
  group_by(date, tank_treatment) %>%
  summarize(
    ph_mean = mean(value, na.rm = TRUE),
    ph_se = sd(value, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )
  mutate(Treatment= case_when( tank_treatment == "Tank_4_5" ~ "Control",
                                    tank_treatment == "Tank_6_7" ~ "High Heat",
                                    tank_treatment == "Tank_2_3" ~ "Low pH",
                                    tank_treatment == "1 Acclimation" ~ "Acclimation"))


```

## Test for differences

Use the daily_calc df created earlier.

Test for differences between treatments.

```{r}
daily_calc%>%
  aov(flow.L.min~treatment, data=.)%>%
  summary()
 
daily_calc%>%
  aov(par.avg~treatment, data=.)%>%
  summary()

daily_calc%>%
  aov(pH.total~treatment, data=.)%>%
  summary()

daily_calc%>%
  aov(pH.nbs~treatment, data=.)%>%
  summary()

daily_calc%>%
  aov(sal.psu~treatment, data=.)%>%
  summary()

daily_calc%>%
  aov(temp.C~treatment, data=.)%>%
  summary()
```

Test for differences between tanks and treatment

```{r}
daily_calc%>%
  aov(flow.L.min~tank*treatment, data=.)
 
daily_calc%>%
  aov(par~tank*treatment, data=.)%>%
  summary()

daily_calc%>%
  aov(pH.total~tank*treatment, data=.)%>%
  summary()

daily_calc%>%
  aov(pH.nbs~tank*treatment, data=.)%>%
  summary()

daily_calc%>%
  aov(sal.psu~tank*treatment, data=.)%>%
  summary()

daily_calc%>%
  aov(temp.C~tank*treatment, data=.)%>%
  summary()
```

## Summarize daily measurements

Calculate descriptive statistics

```{r}
summary<-daily_calc%>%
  group_by(tank, treatment)%>%
  select(!date)%>%
  select(!time)%>%
  summarise(across(everything(), list(mean = mean, sd = sd), na.rm = TRUE)); summary

write_csv(summary, "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/environmental/daily_measurements_summary.csv")
```
