---
title: "Hobo data"
author: "Florence Fields"
date: "2024-10-03"
output: html_document
---

```{r setup, include=FALSE}
#Load libraries 

library(dplyr)
library(purrr)
library(readr)
library(ggplot2)

```


```{r setup, include=FALSE}
file_list <- list.files(path = "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/HobologgerData/20240808", pattern = "*.csv", full.names = TRUE)

```

## Including Plots

You can also embed plots, for example:

```{r setup, echo=FALSE}
# Read all CSV files into a list of data frames and remove specific rows
tank5<- read.csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/HobologgerData/20240815/temp_tank5.csv")
tank4<- read.csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/HobologgerData/20240815/temp_tank4.csv")
tank6<- read.csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/HobologgerData/20240815/temp_tank6.csv")
tank7<- read.csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/HobologgerData/20240815/temp_tank7.csv")
tank3<- read.csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/HobologgerData/20240815/temp_tank3.csv")

#Removing the first coloumn, we don't need it


tank4<- tank4[-c(1)]
tank6<- tank6[-c(1)]
tank7<- tank7[-c(1)]
tank5<- tank5[-c(1)]
tank3<- tank3[-c(1)]

#Renaming the colounms 
tank4 <-tank4 %>% rename(Date.time = X)
tank4 <-tank4 %>% rename(Temperature = X.1)

tank6 <-tank6 %>% rename(Date.time = X)
tank6 <-tank6 %>% rename(Temperature = X.1)

tank5 <-tank5 %>% rename(Date.time = X)
tank5 <-tank5 %>% rename(Temperature = X.1)

tank7 <-tank7 %>% rename(Date.time = X)
tank7 <-tank7 %>% rename(Temperature = X.1)

tank3 <-tank3 %>% rename(Date.time = X)
tank3 <-tank3 %>% rename(Temperature = X.1)

#Removing unwanted row
tank4<- tank4[-c(1),]
tank6<- tank6[-c(1),]
tank7<- tank7[-c(1),]
tank5<- tank5[-c(1),]
tank3<- tank3[-c(1),]


#Assigning tank numbers to each file
tank4 <- tank4 %>%
  mutate(tank = rep("tank4", 1991))
tank6 <- tank6 %>%
  mutate(tank = rep("tank6", 2017))
tank7 <- tank7 %>%
  mutate(tank = rep("tank7", 2002))
tank5 <- tank5 %>%
  mutate(tank = rep("tank5", 1990))
tank3 <- tank3 %>%
  mutate(tank = rep("tank3", 1651))


# Combine all data frames into one
tank_list<- list(tank4,tank5,tank6,tank7,tank3)

tank_merge <-Reduce(function(x, y) merge(x, y, all=TRUE), tank_list)
#tank_merge<- tank_merge[-c(4:7)]
head(tank_merge)

write.csv(tank_merge, "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/environmental/Hobo/20240815TempHobo.csv")

```

#Merging all hobo data and plotting it

```{r}
df1 <- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/environmental/Hobo/20240724TempHobo.csv")

df2 <- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/environmental/Hobo/20240808TempHobo.csv")

df3 <- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/environmental/Hobo/20240815TempHobo.csv")
df1<- df1[-c(1)]
df2<- df2[-c(1)]
df3<- df3[-c(1)]

```

```{r}
#Individual date collection Hobo plots 

ggplot(df1, aes(Date.time, Temperature, color= "tank"))+
  geom_line()


ggplot(df1, aes(x = Date.time, y=Temperature, color= tank)) +
  geom_point() +
  facet_wrap(~ tank, scales = "free_y") +
  labs(x = "Time", y = "Temperature °C") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle=-40),legend.position = "top", 
        legend.text = element_text(size = 7),
                                   legend.title = element_text(size = 8))+
  scale_color_manual(values = c("red", "blue", "black", "purple"))

###Maybe separate just take the avg of the day and plot that 
  
```


# List of data frames to merge
df_list <- list(df1, df2, df3)

# Merge all data frames in the list by "Date.time"
df_merge <- Reduce(function(x, y) merge(x, y, by = "Date.time"), df_list)

# View the merged data frame
print(df_mergee)



ggplot(df_mergee, aes(x = Date.time, y = Temperature)) +
  geom_point(aes(color = tank),size=.5) +
  facet_wrap(~ tank, scales = "free_y") +
  labs(x = "Time", y = "Temperature") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = -40), 
        legend.position = "top", 
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8)) +
  scale_color_manual(values = c("red", "blue", "green", "purple", "orange"))

  ggsave(path = "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/environmental/Hobo", filename = "FinalHobotemp.pdf")
#This isn't working, need to fix this 
```

