---
title: "Photosynthesis and respiration rate calculations"
author: "HM Putnam, AS Huffmyer"
edited by: "DM Becker-Polinski"
date: "20220306"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)
```

```{r load_packages}
## install packages if you dont already have them in your library
if ("devtools" %in% rownames(installed.packages()) == 'FALSE') install.packages('devtools') 
if ("segmented" %in% rownames(installed.packages()) == 'FALSE') install.packages('segmented') 
if ("plotrix" %in% rownames(installed.packages()) == 'FALSE') install.packages('plotrix') 
if ("gridExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('gridExtra') 
if ("LoLinR" %in% rownames(installed.packages()) == 'FALSE') install_github('colin-olito/LoLinR') 
if ("lubridate" %in% rownames(installed.packages()) == 'FALSE') install.packages('lubridate') 
if ("chron" %in% rownames(installed.packages()) == 'FALSE') install.packages('chron') 
if ("plyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('plyr') 
if ("dplyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('dplyr') 
if ("phytotools" %in% rownames(installed.packages()) == 'FALSE') install.packages('phytotools') 
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom') 
if ("purrr" %in% rownames(installed.packages()) == 'FALSE') install.packages('purrr') 

#Read in required libraries

library("devtools")
library("ggplot2")
library("segmented")
library("plotrix")
library("gridExtra")
library("LoLinR")
library("lubridate")
library("chron")
library('phytotools')
library("tidyverse")
library("broom")
library("purrr")
library("dplyr")
```

# Import data
```{r import_data}
Data <- read.csv(file = '../output/physiology/pi_curves/PI.curve.rates.csv', row.names = 1)
```


# Define data  

```{r}
#specify data
Data$PAR <- as.numeric(Data$Light_Value)
Data$Pc <- as.numeric(Data$micromol.cm2.h)
```



# Define PI curve function as a nonlinear Least Squares regression of a quadratic fit, test nls fit
Aquatic Photosynthesis, Falkowski   
Pmax = max photosynthesis (AKA Am from Bayesian script)  
alpha = quantum yeild (AKA AQY from Bayesian script)  
I/E = irradiance (AKA PAR from Bayesian script)  
Rd = dark respiration   

Run models 

Using flexible initial values based on input data:  

```{r}
nls_data <- Data %>% 
   group_by(colony_id) %>%
   nest(data = -colony_id) %>%
   mutate(model1 = purrr::map(data, ~ 
                         nls(Pc ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=(max(.$Pc)-min(.$Pc)),  AQY=0.001, Rd=-min(.$Pc))) %>%
                              tidy %>%
                              dplyr::select(term, estimate) %>% 
                              spread(term, estimate))) %>%
  unnest(model1) %>%
  unnest(data) %>%
  group_by(colony_id) %>%
  dplyr::summarise(Am=mean(Am), AQY=mean(AQY), Rd=mean(Rd))%>%
  write_csv(., "output/physiology/pi_curves/PI_curve_pars_NLS.csv")

```
Plot curve over data points.  
```{r}
augmented <- Data %>% 
  nest(data = -colony_id) %>% 
  mutate(
    fit = purrr::map(data, ~ nls(Pc ~ (Am*((AQY*PAR)/(sqrt(Am^2 + (AQY*PAR)^2)))-Rd), data=., start=list(Am=0.7,  AQY=0.001, Rd=.4))),
    augmented = purrr::map(fit, augment),
  ) %>% 
  unnest(augmented)


#all colonies together
PI_curves <- augmented %>%
  group_by(colony_id)%>%
  qplot(PAR, Pc, data = ., geom = 'point', colour = colony_id) +
  geom_line(aes(y=.fitted))+
  theme_classic()+
  theme(legend.position="none");PI_curves


ggsave("output/physiology/pi_curves/PI_curves_plot.pdf")


#view individual plots
by(augmented,augmented$colony_id, function(i) {
  ggplot(i) +
          geom_point(aes(PAR, Pc, group=colony_id)) + 
          geom_line(aes(y=.fitted, x=PAR)) + 
          theme_classic()+
          labs(x = expression(paste('PAR (', mu, "mol photons m"^-2, 's'^-1,")")),
               y = expression(paste('Photosynthetic rate (', mu, "mol cm"^-2, 'h'^-1,")")),
               title = paste0("1_", augmented$colony_id))
})


```










