---
title: "ENCORE Priming Carbonate Chemistry Analysis"
Script editor: "Florence Fields"
Project: ENCORE
output: html_document
---


This script will calculate various carbonate chemistry parameters for the ENCORE Priming Bermuda 2024 experimental dataset. Using scripts from [Sam Gurr](https://github.com/SamGurr/Geoduck_transgen_offspring_OA/blob/master/RAnalysis/Scripts/CarbChem.R) and Emma Strand [script1](https://github.com/hputnam/Acclim_Dynamics/blob/master/Scripts/CarbChem.Rmd) and [script2](https://github.com/emmastrand/HIMB_Bleaching_2018/blob/master/Scripts/CarbChem.R)


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
rm(list=ls()) # removes all prior objects

setwd("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024") #set working directory

knitr::opts_knit$set(root.dir = "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024")

```


##Read in required libraries
```{r, echo = FALSE}

library(tidyverse)
library(multcomp)
library(car)
library(readxl)
library(seacarb)
library(cowplot)
library(dplyr)
library(ggpubr)
library(ggstatsplot)
library(wesanderson)
library(ggsignif)
library(dunn.test)
library(kableExtra)
library(webshot2)
library(magick)
library(lubridate)
library(reshape2)
library(plyr)
```

##Required Data files
#CRM_TA_Data.csv
#daily_measurements_summary.csv
#~/data/environmental/Tris

```{r, echo = FALSE}
##### DISCRETE pH CALCULATIONS #####
## Read in tris calibration information and calculate slope and intercept 
path <-("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/Tris")

file.names<-list.files(path = path, pattern = "csv$") #list all the file names in the folder to get only get the csv files

pH.cals <- data.frame(matrix(NA, nrow=length(file.names), ncol=3, dimnames=list(file.names,c("Date", "Intercept", "Slope")))) #generate a 3 column dataframe with specific column names

for(i in 1:length(file.names)) { # for every file in list start at the first and run this following function
  Calib.Data <-read.table(file.path(path,file.names[i]), header=TRUE, sep=",", na.string="NA", as.is=TRUE) #reads in the data files
  model <-lm(mVTris ~ Ttris, data=Calib.Data) #runs a linear regression of mV as a function of temperature
  coe <- coef(model) #extracts the coeffecients
  summary(model)$r.squared
  plot(Calib.Data$mVTris, Calib.Data$Ttris)
  pH.cals[i,2:3] <- coe #inserts them in the dataframe
  pH.cals[i,1] <- substr(file.names[i],1,8) #stores the file name in the Date column
}
colnames(pH.cals) <- c("Calib.Date",  "Intercept",  "Slope") #rename columns
pH.cals #view data

#constants for use in pH calculation 
R <- 8.31447215 #gas constant in J mol-1 K-1 
F <-96485.339924 #Faraday constant in coulombs mol-1

```

## Read in probe daily measurements
```{r}
daily <- read.csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/environmental/daily_measurements.csv")
colnames(daily)[11] <- 'Calib.Date'

daily <- daily[, -c(9, 10, 12)] # Remove columns by their indices
daily$Calib.Date <- as.factor(daily$Calib.Date)
pH.cals$Calib.Date <- as.factor(pH.cals$Calib.Date)
```

## Merge daily measurements w/ pH tris
```{r}
SW.chem <- merge(pH.cals, daily, by = "Calib.Date")
```

## Calculate total pH
```{r}
mvTris <- SW.chem$temp.C*SW.chem$Slope+SW.chem$Intercept #calculate the mV of the tris standard using the temperature mv relationships in the measured standard curves 
STris<-35 #salinity of the Tris
phTris<- (11911.08-18.2499*STris-0.039336*STris^2)*(1/(SW.chem$temp.C+273.15))-366.27059+ 0.53993607*STris+0.00016329*STris^2+(64.52243-0.084041*STris)*log(SW.chem$temp.C+273.15)-0.11149858*(SW.chem$temp.C+273.15) #calculate the pH of the tris (Dickson A. G., Sabine C. L. and Christian J. R., SOP 6a)
SW.chem$pH.Total<-phTris+(mvTris/1000-SW.chem$pH.mV/1000)/(R*(SW.chem$temp.C+273.15)*log(10)/F) #calculate the pH on the total scale (Dickson A. G., Sabine C. L. and Christian J. R., SOP 6a)
```

# Discrete TA calculations
```{r}
TA1 <- read.csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/Titrator/Cumulative_TA_Output.csv")

```

#####Altering the Orginal TA dataset to fit the script
```{r}
#Removing duplicate sample rows 
TA <- TA1 %>% slice(-c(22, 24, 27, 53))

# Remove row where Name is "Junk"
TA <- TA[TA$DateSampled != "Junk", ]

#Renaming DateSampled to date
TA<- TA %>% 
  rename(date = DateSampled)
print(TA)
#Renaming tank to Tank in SW.chem
SW.chem2<- SW.chem %>%
  rename(Tank = tank)

#Changing "acc" in tank to 1. Acclimation will be referred to as tank 1
TA <- TA %>% mutate(Tank = ifelse(Tank == "acc", "1", Tank))

# Converting the Date column to Date format
TA$date <- ymd(TA$date)

# Format the Date column to MM/DD/YYYY
TA$date <- format(TA$date, "%m/%d/%Y")

#Removing the first column with the TA process date
TA<-TA[, -c(1)]
print(TA)

#Removing Calib.Date in SW.chem dataframe
SW.chem2<-SW.chem2[, -c(1)]
print(SW.chem2)
```

## Examine for outliers 
```{r}
ggplot(TA, aes(x = PrimingTreatment, y = TA)) +
  geom_boxplot()

#identify outliers by Treatment groups and sample rep
outlier.plot <- ggbetweenstats(TA, PrimingTreatment, TA, outlier.tagging = TRUE)
outlier.plot
```
## Remove outliers
```{r}
#set quantile values
q <- c(0.25, 0.75)

# calculate quantile values by Temperature and Treatment groups
Quants <- TA %>%
  group_by(PrimingTreatment) %>%
  summarize(quant25 = quantile(TA, probs = q[1]),
            quant75 = quantile(TA, probs = q[2]),
            IQRbyGroup=IQR(TA))

#Calculate Quantile upper and lower ranges 
Quants$upper <-  Quants$quant75+1.5*Quants$IQRbyGroup # Upper Range  
Quants$lower <- Quants$quant25-1.5*Quants$IQRbyGroup # Lower Range

#Calculate STDev upper and lower ranges 
Quants$upper.stdev <- mean(TA$TA)+(sd(TA$TA)*2) # Upper Range  
Quants$lower.stdev <- mean(TA$TA)-(sd(TA$TA)*2) # Lower Range

#join outlier cutoffs with rate data
TAdata <- left_join(TA, Quants)

#remove outliers from rates
TAdata <- TAdata %>%
   filter(TA < upper) %>%
   filter(TA > lower)

# Plot TA value by treatment 
ggplot(TAdata, aes(x = PrimingTreatment, y = TA)) +
  geom_boxplot()

# Remove outliers from high trt (rows 48 and 30)
TAdata <- TAdata[-c(48,30),]

# Plot TA value by treatment 
ggplot(TAdata, aes(x = PrimingTreatment, y = TA)) +
  geom_boxplot()

#Selecting data columns
TAdata <- TAdata %>% 
  dplyr::select(c("Tank","PrimingTreatment","TA","Salinity", "date"))
 
#Renaming the treatment column in SW.chem2 

SW.chem2 <- SW.chem2 %>% rename(PrimingTreatment = treatment)

# Renaming within colounms
TAdata$PrimingTreatment<- gsub("Heat","heated",TAdata$PrimingTreatment)
TAdata$PrimingTreatment<- gsub("pH","Low pH",TAdata$PrimingTreatment)
SW.chem2$PrimingTreatment<- gsub("low_ph","Low pH",SW.chem2$PrimingTreatment)
SW.chem2$PrimingTreatment<- gsub("heated","High Heat",SW.chem2$PrimingTreatment)
SW.chem2$PrimingTreatment<- gsub("control","Control",SW.chem2$PrimingTreatment)
print(TAdata)
print(SW.chem2)

```

##Formatting the date in SW.chem 
```{r}

# Parse the dates using lubridate
SW.chem2$date <- mdy(SW.chem2$date)

# Format the dates to include leading zeros
SW.chem2$date <- format(SW.chem2$date, "%m/%d/%Y")
head(SW.chem2)
```


## Merge calculated pH and daily measurements with TAdata 
```{r}
#The TA data needed is the salinity and TA from TAdata. Extract data into a new df
#Selecting data columns
TAdata2 <- TAdata %>% 
  dplyr::select(c("Tank","PrimingTreatment","TA","Salinity", "date"))

# Merge seawater chemistry with total alkalinity and salinity
SW.chem3 <- merge(SW.chem2, TAdata2, by=c("date","Tank"), all = TRUE, sort = T)

# View the merged data frame
print(merged_df)

# Remove NAs
SW.chem4 <- na.omit(SW.chem3)
```

## Select specific columns 
```{r}

SW.chem4 <- SW.chem4 %>%
  dplyr::select(date, time, Tank, PrimingTreatment, temp.C, pH.Total, TA, Salinity) %>%
  rename( "Temperature" = "temp.C",
         "PrimingTreatment" = "PrimingTreatment.x" )
```

##Make outlier plots
```{r}
# Temperature
outlier.plot <- ggbetweenstats(SW.chem4, PrimingTreatment, Temperature, outlier.tagging = TRUE)
outlier.plot # outliers present

# pH total
outlier.plot <- ggbetweenstats(SW.chem4, PrimingTreatment, pH.Total, outlier.tagging = TRUE)
outlier.plot # outliers present 

# Salinity
outlier.plot <- ggbetweenstats(SW.chem4, PrimingTreatment, Salinity, outlier.tagging = TRUE)
outlier.plot

# TA
outlier.plot <- ggbetweenstats(SW.chem4, PrimingTreatment, TA, outlier.tagging = TRUE)
outlier.plot
```
##Calculate CO2 parameters
```{r}
carb.output <- seacarb::carb(flag=8, var1=SW.chem4$pH.Total, var2=SW.chem4$TA/1000000, S= SW.chem4$Salinity, T=SW.chem4$Temperature, P=0, Pt=0, Sit=0, pHscale="T", kf="pf", k1k2="l", ks="d") #calculate seawater chemistry parameters using seacarb
carb.output$ALK <- carb.output$ALK*1000000 #convert to µmol kg-1
carb.output$CO2 <- carb.output$CO2*1000000 #convert to µmol kg-1
carb.output$HCO3 <- carb.output$HCO3*1000000 #convert to µmol kg-1
carb.output$CO3 <- carb.output$CO3*1000000 #convert to µmol kg-1
carb.output$DIC <- carb.output$DIC*1000000 #convert to µmol kg-1
carb.output2 <- carb.output[,-c(1,4,5,8,10:13,19)] #subset variables of interest

# Instead of using cbind from previous scripts I will be merging carb.output2 using pH with SW chem info 
carb.output2 <- carb.output2 %>% rename(pH.Total = pH)
carb.output3 <- merge(SW.chem4, carb.output2, by=c("pH.Total"), all = TRUE, sort = T)

# Add a new column based on another column
carb.output3 <- carb.output3 %>%
  mutate(SecondaryStressExposure = case_when(Tank == "2"~"High Heat",
                                            Tank == "3"~"Control",
                                            Tank == "4"~"High Heat",
                                            Tank == "5"~"Control",
                                            Tank == "6"~"High Heat",
                                            Tank == "7"~"Control",
                                            TRUE~NA_character_))

carb.output3<- carb.output3%>%
  dplyr::select(c("date","time","Tank","PrimingTreatment","SecondaryStressExposure","S","T","pH.Total","CO2","pCO2","HCO3","CO3","DIC","ALK","OmegaAragonite"))

colnames(carb.output3) <- c("Date","Time","Tank","PrimingTreatment","SecondaryStressExposure","Salinity","Temperature","pH.Total","CO2","pCO2","HCO3","CO3","DIC","TA","AragSat")

print(carb.output3)

```

#Write output table of parameters + Summarize all output parameters
```{r}
write.csv(carb.output3, paste0("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/environmental/Seawater_Chemistry_Table_Output_All.csv"))

carbo.melted <- melt(carb.output3)
mean.carb.output <-ddply(carbo.melted, .(PrimingTreatment,variable),summarize,
                         N=length(na.omit(value)),
                         mean = (mean(value)),
                         sem = (sd)(value)/sqrt(N))
```

```{r}
# pdf("~/MyProjects/Geoduck_Conditioning/RAnalysis/Output/Water_Chem_withTA.pdf")
# CHECK YOUR Y AXIS!
pdf("C:/Users/samjg/Documents/My_Projects/Geoduck_transgen_offspring_OA/RAnalysis/Output/Water_Chem_withTA.pdf")
par(mfrow=c(3,2))
plot(carb.output$Treatment, carb.output$Temperature, xlab="Treatment", ylab="Temperature°C", ylim=c(8,16))
plot(carb.output$Treatment, carb.output$pH, xlab="Treatment", ylab="pH Total Scale", ylim=c(6.5,8.2))
plot(carb.output$Treatment, carb.output$pCO2, xlab="Treatment", ylab="pCO2 µmol kg-1", ylim=c(350,8000))
plot(carb.output$Treatment, carb.output$Salinity, xlab="Treatment", ylab="Salinity psu", ylim=c(25,30))
plot(carb.output$Treatment, carb.output$TA, xlab="Treatment", ylab="Total Alkalinity µmol kg-1", ylim=c(1800,2400))
plot(carb.output$Treatment, carb.output$Aragonite.Sat, xlab="Treatment", ylab="Aragonite Saturation State", ylim=c(0,2))
dev.off()

```

#Run Anova for each parameter
```{r}
## Test for differences in salinity by treatment with one-way ANOVA
res.aov <- aov(Salinity ~ PrimingTreatment, data = carb.output3)
summary(res.aov)
TukeyHSD(res.aov) # See if there is any differences between specific treatments 
plot(res.aov, 1) # Check ANOVA assumptions of normality 
leveneTest(Salinity ~ PrimingTreatment, data = carb.output3) # Check ANOVA assumptions of homogenity of variance 

## Test for differences in temperature by treatment with one-way ANOVA
res.aov <- aov(Temperature ~ Treatment, data = carb.output)
summary(res.aov)
TukeyHSD(res.aov) # See if there is any differences between specific treatments 
plot(res.aov, 1) # Check ANOVA assumptions of normality 
leveneTest(Temperature ~ Treatment, data = carb.output) # Check ANOVA assumptions of homogenity of variance 

## Test for differences in pH by treatment with one-way ANOVA
res.aov <- aov(pH.Total ~ Treatment, data = carb.output)
summary(res.aov)
TukeyHSD(res.aov) # See if there is any differences between specific treatments 
plot(res.aov, 1) # Check ANOVA assumptions of normality 
leveneTest(pH.Total ~ Treatment, data = carb.output) # Check ANOVA assumptions of homogenity of variance 
## Test for differences in CO2 by treatment with one-way ANOVA
res.aov <- aov(CO2 ~ Treatment, data = carb.output)
summary(res.aov)
TukeyHSD(res.aov) # See if there is any differences between specific treatments 
plot(res.aov, 1) # Check ANOVA assumptions of normality 
leveneTest(CO2 ~ Treatment, data = carb.output) # Check ANOVA assumptions of homogenity of variance --NOT HOMOGENOUS
```




