---
title: "Photosynthesis and respiration rate calculations"
author: "Hollie Putnam "
edited: "Florence Fields"
date: "2024-06-16"
output: html_document
---

This script was adapted and edited from both Hollie Putnam and Danielle Becker's P&R script. This edited script incorperates run and sample metadata, P&R files and the surface area
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#clear lists
rm(list=ls())

## install packages if you dont already have them in your library
if (!require("devtools")) install.packages("devtools")
if (!require("furrr")) install.packages("furrr")
if (!require("future")) install.packages("future")
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("gridExtra")) install.packages("gridExtra")
if (!require("ggpubr")) install.packages("ggpubr")
if (!require("lubridate")) install.packages("lubridate")
if (!require("cowplot")) install.packages("cowplot")
if (!require("LoLinR")) install_github('colin-olito/LoLinR') 
if (!require("ggstatsplot")) install.packages("ggstatsplot")
if ("rTPC" %in% rownames(installed.packages()) == 'FALSE') remotes::install_github("padpadpadpad/rTPC")
if ("nls.multstart" %in% rownames(installed.packages()) == 'FALSE') install.packages('nls.multstart')
if ("broom" %in% rownames(installed.packages()) == 'FALSE') install.packages('broom') 

#Read in required libraries
library(devtools)
library(LoLinR)
library(tidyverse)
library(gridExtra)
library(ggpubr)
library(lubridate)
library(cowplot)
library(ggstatsplot)
library(rTPC)
library(nls.multstart)
library(broom)
library(car)
library(scales)

## libraries for parallel processing
library(future)
library(furrr)
library(purrr)
```

#Week 1 of PR- July 08 -July 09 2024

## Set working directory and get the file path
```{r, warning = FALSE}

setwd("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/PR_use/")

path.p<-"C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/PR_use/20240708_09"

getwd()
```

### Import data files
```{r, warning = FALSE}
#list all csv file names
file.names<-basename(list.files(path = path.p, pattern = "csv$", recursive = TRUE))
# omit metadata from files to be read in as data
file.names <- file.names[!grepl("metadata", file.names)]   

# Load Surface area
SA <- read.csv(file="C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/Surface_Area/Surface_Area.csv", header=T)

# Load sample metadata
Sample.Info <- read.csv(file="C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/PR_use/20240708_09/pr_sample_metadata.csv", header=T) #read in sample.info data

# Load run metadata
run.info <- read_csv(file = "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/PR_use/20240708_09/pr_run_metadata.csv")
```

##Formating dataset for script
```{r, warning = FALSE}
# Join run and sample metadta first by sample id

colnames(SA)[colnames(SA) == "Colony_id"] <- "sample.id" #changing the column name in SA to match the sample metadata column name
metadata <- full_join(Sample.Info, run.info)%>%
  mutate(Date = as_date(as.character(Date), format = "%Y%m%d", tz = ""))


# Select only certain columnns
metadata <- metadata %>%
  select(sample.id, Run, Treatment, Species, Chamber.Vol.L, Date, Start.time, Stop.time, Light_Percent, Temp.Cat)

# Read in all data files
#edit sample.id name to remove .csv and the date from files
df <- tibble(file.name = file.names) %>%
  mutate(sample.id = gsub("_.*", "", file.name))%>%  # Get sample from filename and select for info before the first _
  mutate(sample.id = gsub(".csv", "", sample.id),  #remove .csv from sample.id
          info = map(sample.id, ~filter(metadata, sample.id == .)),           # Get associated sample info
         data0 = map(file.name, ~read_csv(file.path(path.p, .), skip = 1, col_types = cols(.default = "d", Time = "t"))))   # Get associated O2 data

# Select only Time, Value, and Temp columns from O2 data
df <- df %>%
  mutate(data0 = map(data0, ~select(., Time, Value, Temp)))%>%
  mutate(data0 = map(data0, ~(.x %>% filter(complete.cases(.))))) #remove NAs 

```

## Use the time breaks in the sample info to link O2 data with light levels
```{r, warning = FALSE}
df <- df %>%
  mutate(intervals = map2(data0, info, function(.x, .y) {
    split(.x, f = cut(as.numeric(.x$Time), breaks = as.numeric(c(.y$Start.time, last(.y$Stop.time))),
                      labels = as.character(.y$Light_Percent)))})) %>%
  mutate(data = map(intervals, ~ unnest(tibble(.), .id = "Light_Percent")))

## 'data' now contains the O2 data with the corresponding light level as another column
## Example of what 'data' for each sample looks like:
# df$data[[1]]

```


## Thin data
```{r, fig.height = 8, fig.width = 8}
# Set thinning parameter
thin_par <- 20

# Thin data for all samples
df <- df %>%
  mutate(thin_data = map(data, ~ slice(., seq(1, nrow(.), thin_par))))


# Create plots for full dataset and thinned data
df <- df %>%
  mutate(data_plot = map2(data, sample.id, ~ ggplot(.x, aes(x = Time, y = Value)) + 
                            facet_wrap(~ as.numeric(Light_Percent), scales = "free") +
                            geom_point() +
                            labs(title = .y)),
    thin_data_plot = map2(thin_data, sample.id, ~ ggplot(.x, aes(x = Time, y = Value)) + 
                            facet_wrap(~ as.numeric(Light_Percent), scales = "free") +
                            geom_point() +
                            labs(title = .y)))

# Example of plots
cowplot::plot_grid(df$data_plot[[1]], df$thin_data_plot[[1]], nrow = 2,
                   labels = c("Example plot: all data", "Example plot: thinned data"))



```

## The full or thinned data plot for any sample can be accessed like this:
```{r, warning = FALSE}
df %>%
  filter(sample.id == "DL-1-11") %>%
  pull(thin_data_plot)
```

# Fit regressions to each interval for each sample
```{r, warning = FALSE}
# Define function for fitting LoLinR regressions to be applied to all intervals for all samples
fit_reg <- function(df) {
  rankLocReg(xall = as.numeric(df$Time), yall = df$Value, 
             alpha = 0.2, method = "pc", verbose = FALSE)
}

# Setup for parallel processing
plan(multisession)

# Map LoLinR function onto all intervals of each sample's thinned dataset
df <- df %>%
  mutate(regs = furrr::future_map(thin_data, function(.) {       # future_map executes function in parallel
    group_by(., Light_Percent) %>%
    do(rankLcRg = fit_reg(.))
  }))

## Now 'regs' contains the fitted local regressions for each interval of each sample's thinned dataset

# Define function to pull out and plot regression diagnostics
plot_rankLcRg <- function(sample.id, interval_number) {
  df %>%
    filter(sample.id == sample.id) %>%
    pluck("regs", 1, "rankLcRg", interval_number) %>%
    plot()
}
```
## Extract slope of best regression for each interval for each sample
```{r}
df.out <- df %>% 
  unnest(regs) %>%
  mutate(micromol.L.s = map_dbl(rankLcRg, ~ pluck(., "allRegs", "b1", 1)))
```

## Adjust by chamber volume and normalize to surface area
```{r}
### Merge rates with sample info
pr <- left_join(
  select(df.out, sample.id, Light_Percent, micromol.L.s),
  distinct(metadata, sample.id, Treatment, Run, Chamber.Vol.L, Date)
)

## Correct for chamber volume and blanks
pr <- pr %>%
  mutate(micromol.s = micromol.L.s * Chamber.Vol.L)

# Get blank values -- average for each run and light value in case multiple blanks
blanks <- pr %>%
  filter(grepl("BLANK", sample.id)) %>%
  group_by(Run, Light_Percent) %>%
  summarise(micromol.s.blank = mean(micromol.s))

# Join blank values with rest of data and subtract values from samples for same run and light value
pr <- left_join(pr, blanks) %>%
  mutate(micromol.s.adj = micromol.s - micromol.s.blank) %>%
  # After correcting for blank values, remove blanks from data
  filter(!grepl("BLANK", sample.id))

# Join surface area with rest of data
pr <- left_join(pr, select(SA, sample.id, surface.area.cm2))

# Normalize rates by surface area
pr <- pr %>%
  mutate(micromol.cm2.s = micromol.s.adj / surface.area.cm2,
         micromol.cm2.h = micromol.cm2.s * 3600)

Photo <- pr %>%
  filter(Light_Percent ==331)%>% 
  separate(sample.id, c("Species","Genotype","Rep"), sep = "-", remove=FALSE)

Resp <- pr %>%
  filter(Light_Percent ==0)%>% 
  separate(sample.id, c("Species","Genotype","Rep"), sep = "-", remove=FALSE)

Photo <- na.omit(Photo) %>% select("sample.id", micromol.cm2.h, Treatment, Genotype, Species)
#colnames(Photo)[1] <- "sample.id"
Photo$Timepoint <- "Week1"
Photo$measure <- "Photo"
Resp <- na.omit(Resp) %>% select("sample.id", micromol.cm2.h, Treatment, Genotype, Species)
#colnames(Resp)[1] <- "sample.id"
Resp$Timepoint <- "Week1"
Resp$measure <- "Resp"

write.csv(Photo, file="C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Photo_Rawextracted_rates_Week1.csv")
write.csv(Resp, file="C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Resp_Rawextracted_rates_Week1.csv")
```

```{r}

df_summary_Resp <- Resp %>%
  group_by(Genotype, Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))

df_mean_Resp <- Resp %>%
  group_by(Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))


# Visualize data
pdf("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Raw_Resp_Week1.pdf", width=6, height=6)
r_plot<-df_summary_Resp %>%
    ggplot(., aes(x = Timepoint, y = -mean_micromol_cm2_h, colour=Treatment)) +
    geom_point(aes(fill=Treatment, group=Genotype, shape=Genotype), size=2) + 
    geom_point(data = df_mean_Resp, aes(x = Timepoint, y = -mean_micromol_cm2_h, colour=Treatment), size = 4) + 
    xlab("Timepoint") +
    facet_wrap(vars(Species))+
    scale_fill_manual(name="Treatment", values=c("gray","red","blue"))+
    scale_color_manual(name="Treatment", values=c("gray","red","blue"))+
    ylab(expression(bold(paste("Resp rate (µmol ", O[2], " cm"^-2, "min"^-1, ")")))) +
    #ylim(0,3)+
    theme_classic() + 
    theme(axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)); r_plot
if (dev.cur() > 1) {
  dev.off()
}

df_summary_Photo <- Photo %>%
  group_by(Genotype, Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))

df_mean_Photo <- Photo %>%
  group_by(Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))


# Visualize data
pdf("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Raw_Photo_Week1.pdf", width=6, height=6)
r_plot<-df_summary_Photo %>%
    ggplot(., aes(x = Timepoint, y = mean_micromol_cm2_h, colour=Treatment)) +
    geom_point(aes(fill=Treatment, group=Genotype, shape=Genotype), size=2) + 
    geom_point(data = df_mean_Photo, aes(x = Timepoint, y = mean_micromol_cm2_h, colour=Treatment), size = 4) + 
    xlab("Timepoint") +
    facet_wrap(vars(Species))+
    scale_fill_manual(name="Treatment", values=c("gray","red","blue"))+
    scale_color_manual(name="Treatment", values=c("gray","red","blue"))+
    ylab(expression(bold(paste("Photo rate (µmol ", O[2], " cm"^-2, "min"^-1, ")")))) +
    #ylim(0,3)+
    theme_classic() + 
    theme(axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)); r_plot
if (dev.cur() > 1) {
  dev.off()
}

```

#Week 2 of PR- July 16 -July 17 2024
## Set working directory and get the file path
```{r, warning = FALSE}

setwd("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/PR_use/")

path.p<-"C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/PR_use/20240716_17"

getwd()
```

### Import data files
```{r, warning = FALSE}
#list all csv file names
file.names<-basename(list.files(path = path.p, pattern = "csv$", recursive = TRUE))
# omit metadata from files to be read in as data
file.names <- file.names[!grepl("metadata", file.names)]   

# Load Surface area
SA <- read.csv(file="C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/Surface_Area/Surface_Area.csv", header=T)

# Load sample metadata
Sample.Info <- read.csv(file="C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/PR_use/20240716_17/pr_sample_metadata.csv", header=T) #read in sample.info data

# Load run metadata
run.info <- read_csv(file = "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/PR_use/20240716_17/pr_run_metadata.csv")
```

##Formating dataset for script
```{r, warning = FALSE}
# Join run and sample metadta first by sample id

colnames(SA)[colnames(SA) == "Colony_id"] <- "sample.id" #changing the column name in SA to match the sample metadata column name
metadata <- full_join(Sample.Info, run.info)%>%
  mutate(Date = as_date(as.character(Date), format = "%Y%m%d", tz = ""))


# Select only certain columnns
metadata <- metadata %>%
  select(sample.id, Run, Treatment, Species, Chamber.Vol.L, Date, Start.time, Stop.time, Light_Percent, Temp.Cat)

# Read in all data files
#edit sample.id name to remove .csv and the date from files
df <- tibble(file.name = file.names) %>%
  mutate(sample.id = gsub("_.*", "", file.name))%>%  # Get sample from filename and select for info before the first _
  mutate(sample.id = gsub(".csv", "", sample.id),  #remove .csv from sample.id
          info = map(sample.id, ~filter(metadata, sample.id == .)),           # Get associated sample info
         data0 = map(file.name, ~read_csv(file.path(path.p, .), skip = 1, col_types = cols(.default = "d", Time = "t"))))   # Get associated O2 data

# Select only Time, Value, and Temp columns from O2 data
df <- df %>%
  mutate(data0 = map(data0, ~select(., Time, Value, Temp)))%>%
  mutate(data0 = map(data0, ~(.x %>% filter(complete.cases(.))))) #remove NAs 

```

## Use the time breaks in the sample info to link O2 data with light levels
```{r, warning = FALSE}
df <- df %>%
  mutate(intervals = map2(data0, info, function(.x, .y) {
    split(.x, f = cut(as.numeric(.x$Time), breaks = as.numeric(c(.y$Start.time, last(.y$Stop.time))),
                      labels = as.character(.y$Light_Percent)))})) %>%
  mutate(data = map(intervals, ~ unnest(tibble(.), .id = "Light_Percent")))

## 'data' now contains the O2 data with the corresponding light level as another column
## Example of what 'data' for each sample looks like:
# df$data[[1]]

```


## Thin data
```{r, fig.height = 8, fig.width = 8}
# Set thinning parameter
thin_par <- 20

# Thin data for all samples
df <- df %>%
  mutate(thin_data = map(data, ~ slice(., seq(1, nrow(.), thin_par))))


# Create plots for full dataset and thinned data
df <- df %>%
  mutate(data_plot = map2(data, sample.id, ~ ggplot(.x, aes(x = Time, y = Value)) + 
                            facet_wrap(~ as.numeric(Light_Percent), scales = "free") +
                            geom_point() +
                            labs(title = .y)),
    thin_data_plot = map2(thin_data, sample.id, ~ ggplot(.x, aes(x = Time, y = Value)) + 
                            facet_wrap(~ as.numeric(Light_Percent), scales = "free") +
                            geom_point() +
                            labs(title = .y)))

# Example of plots
cowplot::plot_grid(df$data_plot[[1]], df$thin_data_plot[[1]], nrow = 2,
                   labels = c("Example plot: all data", "Example plot: thinned data"))



```

## The full or thinned data plot for any sample can be accessed like this:
```{r, warning = FALSE}
df %>%
  filter(sample.id == "DL-2-18") %>%
  pull(thin_data_plot)
```

# Fit regressions to each interval for each sample
```{r, warning = FALSE}
# Define function for fitting LoLinR regressions to be applied to all intervals for all samples
fit_reg <- function(df) {
  rankLocReg(xall = as.numeric(df$Time), yall = df$Value, 
             alpha = 0.2, method = "pc", verbose = FALSE)
}

# Setup for parallel processing
plan(multisession)

# Map LoLinR function onto all intervals of each sample's thinned dataset
df <- df %>%
  mutate(regs = furrr::future_map(thin_data, function(.) {       # future_map executes function in parallel
    group_by(., Light_Percent) %>%
    do(rankLcRg = fit_reg(.))
  }))


## Now 'regs' contains the fitted local regressions for each interval of each sample's thinned dataset

# Define function to pull out and plot regression diagnostics
plot_rankLcRg <- function(sample.id, interval_number) {
  df1 %>%
    filter(sample.id == sample.id) %>%
    pluck("regs", 1, "rankLcRg", interval_number) %>%
    plot()
}
```
## Extract slope of best regression for each interval for each sample
```{r}
df.out <- df %>% 
  unnest(regs) %>%
  mutate(micromol.L.s = map_dbl(rankLcRg, ~ pluck(., "allRegs", "b1", 1)))
```

## Adjust by chamber volume and normalize to surface area
```{r}
### Merge rates with sample info
pr <- left_join(
  select(df.out, sample.id, Light_Percent, micromol.L.s),
  distinct(metadata, sample.id, Treatment, Run, Chamber.Vol.L, Date)
)

## Correct for chamber volume and blanks
pr <- pr %>%
  mutate(micromol.s = micromol.L.s * Chamber.Vol.L)

# Get blank values -- average for each run and light value in case multiple blanks
blanks <- pr %>%
  filter(grepl("BLANK", sample.id)) %>%
  group_by(Run, Light_Percent) %>%
  summarise(micromol.s.blank = mean(micromol.s))

# Join blank values with rest of data and subtract values from samples for same run and light value
pr <- left_join(pr, blanks) %>%
  mutate(micromol.s.adj = micromol.s - micromol.s.blank) %>%
  # After correcting for blank values, remove blanks from data
  filter(!grepl("BLANK", sample.id))

# Join surface area with rest of data
pr <- left_join(pr, select(SA, sample.id, surface.area.cm2))

# Normalize rates by surface area
pr <- pr %>%
  mutate(micromol.cm2.s = micromol.s.adj / surface.area.cm2,
         micromol.cm2.h = micromol.cm2.s * 3600)

Photo <- pr %>%
  filter(Light_Percent ==331)%>% 
  separate(sample.id, c("Species","Genotype","Rep"), sep = "-", remove=FALSE)

Resp <- pr %>%
  filter(Light_Percent ==0)%>% 
  separate(sample.id, c("Species","Genotype","Rep"), sep = "-", remove=FALSE)

Photo <- na.omit(Photo) %>% select("sample.id", micromol.cm2.h, Treatment, Genotype, Species)
#colnames(Photo)[1] <- "sample.id"
Photo$Timepoint <- "Week2"
Photo$measure <- "Photo"
Resp <- na.omit(Resp) %>% select("sample.id", micromol.cm2.h, Treatment, Genotype, Species)
#colnames(Resp)[1] <- "sample.id"
Resp$Timepoint <- "Week2"
Resp$measure <- "Resp"

write.csv(Photo, file="C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Photo_Rawextracted_rates_Week2.csv")
write.csv(Resp, file="C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Resp_Rawextracted_rates_Week2.csv")
```

```{r}

df_summary_Resp <- Resp %>%
  group_by(Genotype, Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))

df_mean_Resp <- Resp %>%
  group_by(Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))


# Visualize data
pdf("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Raw_Resp_Week1.pdf", width=6, height=6)
r_plot<-df_summary_Resp %>%
    ggplot(., aes(x = Timepoint, y = -mean_micromol_cm2_h, colour=Treatment)) +
    geom_point(aes(fill=Treatment, group=Genotype, shape=Genotype), size=2) + 
    geom_point(data = df_mean_Resp, aes(x = Timepoint, y = -mean_micromol_cm2_h, colour=Treatment), size = 4) + 
    xlab("Timepoint") +
    facet_wrap(vars(Species))+
    scale_fill_manual(name="Treatment", values=c("gray","red","blue"))+
    scale_color_manual(name="Treatment", values=c("gray","red","blue"))+
    ylab(expression(bold(paste("Resp rate (µmol ", O[2], " cm"^-2, "min"^-1, ")")))) +
    #ylim(0,3)+
    theme_classic() + 
    theme(axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)); r_plot
if (dev.cur() > 1) {
  dev.off()
}

df_summary_Photo <- Photo %>%
  group_by(Genotype, Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))

df_mean_Photo <- Photo %>%
  group_by(Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))


# Visualize data
pdf("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Raw_Photo_Week1.pdf", width=6, height=6)
r_plot<-df_summary_Photo %>%
    ggplot(., aes(x = Timepoint, y = mean_micromol_cm2_h, colour=Treatment)) +
    geom_point(aes(fill=Treatment, group=Genotype, shape=Genotype), size=2) + 
    geom_point(data = df_mean_Photo, aes(x = Timepoint, y = mean_micromol_cm2_h, colour=Treatment), size = 4) + 
    xlab("Timepoint") +
    facet_wrap(vars(Species))+
    scale_fill_manual(name="Treatment", values=c("gray","red","blue"))+
    scale_color_manual(name="Treatment", values=c("gray","red","blue"))+
    ylab(expression(bold(paste("Photo rate (µmol ", O[2], " cm"^-2, "min"^-1, ")")))) +
    #ylim(0,3)+
    theme_classic() + 
    theme(axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)); r_plot
if (dev.cur() > 1) {
  dev.off()
}

```


#Week 3 of PR- July22 - July23 2024

## Set working directory and get the file path
```{r, warning = FALSE}

setwd("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/PR_use")

path.p<-"C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/PR_use/20240722_23"

getwd()
```

## Import data files
```{r, warning = FALSE}
#list all csv file names
file.names<-basename(list.files(path = path.p, pattern = "csv$", recursive = TRUE))
# omit metadata from files to be read in as data
file.names <- file.names[!grepl("metadata", file.names)]   

# Load Surface area
SA <- read.csv(file="C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/Surface_Area/Surface_Area.csv", header=T)

# Load sample metadata
Sample.Info <- read.csv(file="C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/PR_use/20240722_23/pr_sample_metadata.csv", header=T) #read in sample.info data

# Load run metadata
run.info <- read_csv(file = "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/PR_use/20240722_23/pr_run_metadata.csv")
```

#Formating dataset for script
```{r, warning = FALSE}
# Join run and sample metadta first by sample id

colnames(SA)[colnames(SA) == "Colony_id"] <- "sample.id" #changing the column name in SA to match the sample metadata column name
metadata <- full_join(Sample.Info, run.info)%>%
  mutate(Date = as_date(as.character(Date), format = "%Y%m%d", tz = ""))


# Select only certain columnns
metadata <- metadata %>%
  select(sample.id, Run, Treatment, Species, Chamber.Vol.L, Date, Start.time, Stop.time, Light_Percent, Temp.Cat)

# Read in all data files
#edit sample.id name to remove .csv and the date from files
df <- tibble(file.name = file.names) %>%
  mutate(sample.id = gsub("_.*", "", file.name))%>%  # Get sample from filename and select for info before the first _
  mutate(sample.id = gsub(".csv", "", sample.id),  #remove .csv from sample.id
          info = map(sample.id, ~filter(metadata, sample.id == .)),           # Get associated sample info
         data0 = map(file.name, ~read_csv(file.path(path.p, .), skip = 1, col_types = cols(.default = "d", Time = "t"))))   # Get associated O2 data

# Select only Time, Value, and Temp columns from O2 data
df <- df %>%
  mutate(data0 = map(data0, ~select(., Time, Value, Temp)))%>%
  mutate(data0 = map(data0, ~(.x %>% filter(complete.cases(.))))) #remove NAs 

```

## Use the time breaks in the sample info to link O2 data with light levels
```{r, warning = FALSE}
df <- df %>%
  mutate(intervals = map2(data0, info, function(.x, .y) {
    split(.x, f = cut(as.numeric(.x$Time), breaks = as.numeric(c(.y$Start.time, last(.y$Stop.time))),
                      labels = as.character(.y$Light_Percent)))})) %>%
  mutate(data = map(intervals, ~ unnest(tibble(.), .id = "Light_Percent")))

## 'data' now contains the O2 data with the corresponding light level as another column
## Example of what 'data' for each sample looks like:
# df$data[[1]]

```


## Thin data
```{r, fig.height = 8, fig.width = 8}
# Set thinning parameter
thin_par <- 20

# Thin data for all samples
df <- df %>%
  mutate(thin_data = map(data, ~ slice(., seq(1, nrow(.), thin_par))))

# Create plots for full dataset and thinned data
df <- df %>%
  mutate(data_plot = map2(data, sample.id, ~ ggplot(.x, aes(x = Time, y = Value)) + 
                            facet_wrap(~ as.numeric(Light_Percent), scales = "free") +
                            geom_point() +
                            labs(title = .y)),
    thin_data_plot = map2(thin_data, sample.id, ~ ggplot(.x, aes(x = Time, y = Value)) + 
                            facet_wrap(~ as.numeric(Light_Percent), scales = "free") +
                            geom_point() +
                            labs(title = .y)))

# Example of plots
cowplot::plot_grid(df$data_plot[[1]], df$thin_data_plot[[1]], nrow = 2,
                   labels = c("Example plot: all data", "Example plot: thinned data"))



```

## The full or thinned data plot for any sample can be accessed like this:
```{r, warning = FALSE}
df %>%
  filter(sample.id == "MD-1-12") %>%
  pull(thin_data_plot)
```

# Fit regressions to each interval for each sample
```{r, warning = FALSE}
# Define function for fitting LoLinR regressions to be applied to all intervals for all samples
fit_reg <- function(df) {
  rankLocReg(xall = as.numeric(df$Time), yall = df$Value, 
             alpha = 0.2, method = "pc", verbose = FALSE)
}

# Setup for parallel processing
plan(multisession)

# Map LoLinR function onto all intervals of each sample's thinned dataset
df <- df %>%
  mutate(regs = furrr::future_map(thin_data, function(.) {       # future_map executes function in parallel
    group_by(., Light_Percent) %>%
    do(rankLcRg = fit_reg(.))
  }))

## Now 'regs' contains the fitted local regressions for each interval of each sample's thinned dataset

# Define function to pull out and plot regression diagnostics
plot_rankLcRg <- function(sample.id, interval_number) {
  df %>%
    filter(sample.id == sample.id) %>%
    pluck("regs", 1, "rankLcRg", interval_number) %>%
    plot()
}
```
## Extract slope of best regression for each interval for each sample
```{r}
df.out <- df %>% 
  unnest(regs) %>%
  mutate(micromol.L.s = map_dbl(rankLcRg, ~ pluck(., "allRegs", "b1", 1)))
```

## Adjust by chamber volume and normalize to surface area
```{r}
### Merge rates with sample info
pr <- left_join(
  select(df.out, sample.id, Light_Percent, micromol.L.s),
  distinct(metadata, sample.id, Treatment, Run, Chamber.Vol.L, Date)
)

## Correct for chamber volume and blanks
pr <- pr %>%
  mutate(micromol.s = micromol.L.s * Chamber.Vol.L)

# Get blank values -- average for each run and light value in case multiple blanks
blanks <- pr %>%
  filter(grepl("BLANK", sample.id)) %>%
  group_by(Run, Light_Percent) %>%
  summarise(micromol.s.blank = mean(micromol.s))

# Join blank values with rest of data and subtract values from samples for same run and light value
pr <- left_join(pr, blanks) %>%
  mutate(micromol.s.adj = micromol.s - micromol.s.blank) %>%
  # After correcting for blank values, remove blanks from data
  filter(!grepl("BLANK", sample.id))

# Join surface area with rest of data
pr <- left_join(pr, select(SA, sample.id, surface.area.cm2))

# Normalize rates by surface area
pr <- pr %>%
  mutate(micromol.cm2.s = micromol.s.adj / surface.area.cm2,
         micromol.cm2.h = micromol.cm2.s * 3600)

Photo <- pr %>%
  filter(Light_Percent ==331)%>% 
  separate(sample.id, c("Species","Genotype","Rep"), sep = "-", remove=FALSE)

Resp <- pr %>%
  filter(Light_Percent ==0)%>% 
  separate(sample.id, c("Species","Genotype","Rep"), sep = "-", remove=FALSE)

Photo <- na.omit(Photo) %>% select("sample.id", micromol.cm2.h, Treatment, Genotype, Species)
#colnames(Photo)[1] <- "sample.id"
Photo$Timepoint <- "Week3"
Photo$measure <- "Photo"
Resp <- na.omit(Resp) %>% select("sample.id", micromol.cm2.h, Treatment, Genotype, Species)
#colnames(Resp)[1] <- "sample.id"
Resp$Timepoint <- "Week3"
Resp$measure <- "Resp"

write.csv(Photo, file="C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Photo_Rawextracted_rates_Week3.csv")
write.csv(Resp, file="C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Resp_Rawextracted_rates_Week3.csv")
```

```{r}

df_summary_Resp <- Resp %>%
  group_by(Genotype, Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))

df_mean_Resp <- Resp %>%
  group_by(Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))


# Visualize data
pdf("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Raw_Resp_Week3.pdf", width=6, height=6)
r_plot<-df_summary_Resp %>%
    ggplot(., aes(x = Timepoint, y = -mean_micromol_cm2_h, colour=Treatment)) +
    geom_point(aes(fill=Treatment, group=Genotype, shape=Genotype), size=2) + 
    geom_point(data = df_mean_Resp, aes(x = Timepoint, y = -mean_micromol_cm2_h, colour=Treatment), size = 4) + 
    xlab("Timepoint") +
    facet_wrap(vars(Species))+
    scale_fill_manual(name="Treatment", values=c("gray","red","blue"))+
    scale_color_manual(name="Treatment", values=c("gray","red","blue"))+
    ylab(expression(bold(paste("Resp rate (µmol ", O[2], " cm"^-2, "min"^-1, ")")))) +
    #ylim(0,3)+
    theme_classic() + 
    theme(axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)); r_plot
if (dev.cur() > 1) {
  dev.off()
}

df_summary_Photo <- Photo %>%
  group_by(Genotype, Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))

df_mean_Photo <- Photo %>%
  group_by(Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))


# Visualize data
pdf("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Raw_Photo_Week3.pdf", width=6, height=6)
r_plot<-df_summary_Photo %>%
    ggplot(., aes(x = Timepoint, y = mean_micromol_cm2_h, colour=Treatment)) +
    geom_point(aes(fill=Treatment, group=Genotype, shape=Genotype), size=2) + 
    geom_point(data = df_mean_Photo, aes(x = Timepoint, y = mean_micromol_cm2_h, colour=Treatment), size = 4) + 
    xlab("Timepoint") +
    facet_wrap(vars(Species))+
    scale_fill_manual(name="Treatment", values=c("gray","red","blue"))+
    scale_color_manual(name="Treatment", values=c("gray","red","blue"))+
    ylab(expression(bold(paste("Photo rate (µmol ", O[2], " cm"^-2, "min"^-1, ")")))) +
    #ylim(0,3)+
    theme_classic() + 
    theme(axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)); r_plot
if (dev.cur() > 1) {
  dev.off()
}

```


#Week 4 of PR- August 05 - August 06 2024

## Set working directory and get the file path
```{r, warning = FALSE}

setwd("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/PR_use")

path.p<-"C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/PR_use/20240805_06"

getwd()
```

## Import data files
```{r, warning = FALSE}
#list all csv file names
file.names<-basename(list.files(path = path.p, pattern = "csv$", recursive = TRUE))
# omit metadata from files to be read in as data
file.names <- file.names[!grepl("metadata", file.names)]   

# Load Surface area
SA <- read.csv(file="C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/Surface_Area/Surface_Area.csv", header=T)

# Load sample metadata
Sample.Info <- read.csv(file="C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/PR_use/20240805_06/pr_sample_metadata.csv", header=T) #read in sample.info data

# Load run metadata
run.info <- read_csv(file = "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/PR_use/20240805_06/pr_run_metadata.csv")
```

#Formating dataset for script
```{r, warning = FALSE}
# Join run and sample metadta first by sample id

colnames(SA)[colnames(SA) == "Colony_id"] <- "sample.id" #changing the column name in SA to match the sample metadata column name
metadata <- full_join(Sample.Info, run.info)%>%
  mutate(Date = as_date(as.character(Date), format = "%Y%m%d", tz = ""))


# Select only certain columnns
metadata <- metadata %>%
  select(sample.id, Run, Treatment, Species, Chamber.Vol.L, Date, Start.time, Stop.time, Light_Percent, Temp.Cat)

# Read in all data files
#edit sample.id name to remove .csv and the date from files
df <- tibble(file.name = file.names) %>%
  mutate(sample.id = gsub("_.*", "", file.name))%>%  # Get sample from filename and select for info before the first _
  mutate(sample.id = gsub(".csv", "", sample.id),  #remove .csv from sample.id
          info = map(sample.id, ~filter(metadata, sample.id == .)),           # Get associated sample info
         data0 = map(file.name, ~read_csv(file.path(path.p, .), skip = 1, col_types = cols(.default = "d", Time = "t"))))   # Get associated O2 data

# Select only Time, Value, and Temp columns from O2 data
df <- df %>%
  mutate(data0 = map(data0, ~select(., Time, Value, Temp)))%>%
  mutate(data0 = map(data0, ~(.x %>% filter(complete.cases(.))))) #remove NAs 

```

## Use the time breaks in the sample info to link O2 data with light levels
```{r, warning = FALSE}
df <- df %>%
  mutate(intervals = map2(data0, info, function(.x, .y) {
    split(.x, f = cut(as.numeric(.x$Time), breaks = as.numeric(c(.y$Start.time, last(.y$Stop.time))),
                      labels = as.character(.y$Light_Percent)))})) %>%
  mutate(data = map(intervals, ~ unnest(tibble(.), .id = "Light_Percent")))

## 'data' now contains the O2 data with the corresponding light level as another column
## Example of what 'data' for each sample looks like:
# df$data[[1]]

```


## Thin data
```{r, fig.height = 8, fig.width = 8}
# Set thinning parameter
thin_par <- 20

# Thin data for all samples
df <- df %>%
  mutate(thin_data = map(data, ~ slice(., seq(1, nrow(.), thin_par))))

# Create plots for full dataset and thinned data
df <- df %>%
  mutate(data_plot = map2(data, sample.id, ~ ggplot(.x, aes(x = Time, y = Value)) + 
                            facet_wrap(~ as.numeric(Light_Percent), scales = "free") +
                            geom_point() +
                            labs(title = .y)),
    thin_data_plot = map2(thin_data, sample.id, ~ ggplot(.x, aes(x = Time, y = Value)) + 
                            facet_wrap(~ as.numeric(Light_Percent), scales = "free") +
                            geom_point() +
                            labs(title = .y)))

# Example of plots
cowplot::plot_grid(df$data_plot[[1]], df$thin_data_plot[[1]], nrow = 2,
                   labels = c("Example plot: all data", "Example plot: thinned data"))



```

## The full or thinned data plot for any sample can be accessed like this:
```{r, warning = FALSE}
df %>%
  filter(sample.id == "DL-1-11") %>%
  pull(thin_data_plot)
```

# Fit regressions to each interval for each sample
```{r, warning = FALSE}
# Define function for fitting LoLinR regressions to be applied to all intervals for all samples
fit_reg <- function(df) {
  rankLocReg(xall = as.numeric(df$Time), yall = df$Value, 
             alpha = 0.2, method = "pc", verbose = FALSE)
}

# Setup for parallel processing
plan(multisession)

# Map LoLinR function onto all intervals of each sample's thinned dataset
df <- df %>%
  mutate(regs = furrr::future_map(thin_data, function(.) {       # future_map executes function in parallel
    group_by(., Light_Percent) %>%
    do(rankLcRg = fit_reg(.))
  }))

## Now 'regs' contains the fitted local regressions for each interval of each sample's thinned dataset

# Define function to pull out and plot regression diagnostics
plot_rankLcRg <- function(sample.id, interval_number) {
  df %>%
    filter(sample.id == sample.id) %>%
    pluck("regs", 1, "rankLcRg", interval_number) %>%
    plot()
}
```
## Extract slope of best regression for each interval for each sample
```{r}
df.out <- df %>% 
  unnest(regs) %>%
  mutate(micromol.L.s = map_dbl(rankLcRg, ~ pluck(., "allRegs", "b1", 1)))
```

## Adjust by chamber volume and normalize to surface area
```{r}
### Merge rates with sample info
pr <- left_join(
  select(df.out, sample.id, Light_Percent, micromol.L.s),
  distinct(metadata, sample.id, Treatment, Run, Chamber.Vol.L, Date)
)

## Correct for chamber volume and blanks
pr <- pr %>%
  mutate(micromol.s = micromol.L.s * Chamber.Vol.L)

# Get blank values -- average for each run and light value in case multiple blanks
blanks <- pr %>%
  filter(grepl("BLANK", sample.id)) %>%
  group_by(Run, Light_Percent) %>%
  summarise(micromol.s.blank = mean(micromol.s))

# Join blank values with rest of data and subtract values from samples for same run and light value
pr <- left_join(pr, blanks) %>%
  mutate(micromol.s.adj = micromol.s - micromol.s.blank) %>%
  # After correcting for blank values, remove blanks from data
  filter(!grepl("BLANK", sample.id))

# Join surface area with rest of data
pr <- left_join(pr, select(SA, sample.id, surface.area.cm2))

# Normalize rates by surface area
pr <- pr %>%
  mutate(micromol.cm2.s = micromol.s.adj / surface.area.cm2,
         micromol.cm2.h = micromol.cm2.s * 3600)

Photo <- pr %>%
  filter(Light_Percent ==331)%>% 
  separate(sample.id, c("Species","Genotype","Rep"), sep = "-", remove=FALSE)

Resp <- pr %>%
  filter(Light_Percent ==0)%>% 
  separate(sample.id, c("Species","Genotype","Rep"), sep = "-", remove=FALSE)

Photo <- na.omit(Photo) %>% select("sample.id", micromol.cm2.h, Treatment, Genotype, Species)
#colnames(Photo)[1] <- "sample.id"
Photo$Timepoint <- "Week4"
Photo$measure <- "Photo"
Resp <- na.omit(Resp) %>% select("sample.id", micromol.cm2.h, Treatment, Genotype, Species)
#colnames(Resp)[1] <- "sample.id"
Resp$Timepoint <- "Week4"
Resp$measure <- "Resp"

write.csv(Photo, file="C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Photo_Rawextracted_rates_Week4.csv")
write.csv(Resp, file="C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Resp_Rawextracted_rates_Week4.csv")
```

```{r}

df_summary_Resp <- Resp %>%
  group_by(Genotype, Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))

df_mean_Resp <- Resp %>%
  group_by(Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))


# Visualize data
pdf("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Raw_Resp_Week4.pdf", width=6, height=6)
r_plot<-df_summary_Resp %>%
    ggplot(., aes(x = Timepoint, y = -mean_micromol_cm2_h, colour=Treatment)) +
    geom_point(aes(fill=Treatment, group=Genotype, shape=Genotype), size=2) + 
    geom_point(data = df_mean_Resp, aes(x = Timepoint, y = -mean_micromol_cm2_h, colour=Treatment), size = 4) + 
    xlab("Timepoint") +
    facet_wrap(vars(Species))+
    scale_fill_manual(name="Treatment", values=c("gray","red","blue"))+
    scale_color_manual(name="Treatment", values=c("gray","red","blue"))+
    ylab(expression(bold(paste("Resp rate (µmol ", O[2], " cm"^-2, "min"^-1, ")")))) +
    #ylim(0,3)+
    theme_classic() + 
    theme(axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)); r_plot
if (dev.cur() > 1) {
  dev.off()
}

df_summary_Photo <- Photo %>%
  group_by(Genotype, Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))

df_mean_Photo <- Photo %>%
  group_by(Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))


# Visualize data
pdf("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Raw_Photo_Week4.pdf", width=6, height=6)
r_plot<-df_summary_Photo %>%
    ggplot(., aes(x = Timepoint, y = mean_micromol_cm2_h, colour=Treatment)) +
    geom_point(aes(fill=Treatment, group=Genotype, shape=Genotype), size=2) + 
    geom_point(data = df_mean_Photo, aes(x = Timepoint, y = mean_micromol_cm2_h, colour=Treatment), size = 4) + 
    xlab("Timepoint") +
    facet_wrap(vars(Species))+
    scale_fill_manual(name="Treatment", values=c("gray","red","blue"))+
    scale_color_manual(name="Treatment", values=c("gray","red","blue"))+
    ylab(expression(bold(paste("Photo rate (µmol ", O[2], " cm"^-2, "min"^-1, ")")))) +
    #ylim(0,3)+
    theme_classic() + 
    theme(axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)); r_plot
if (dev.cur() > 1) {
  dev.off()
}

```

#Week 5 of PR- August 10 - August 11 2024

## Set working directory and get the file path
```{r, warning = FALSE}

setwd("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/PR_use")

path.p<-"C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/PR_use/20240810_11"

getwd()
```

## Import data files
```{r, warning = FALSE}
#list all csv file names
file.names<-basename(list.files(path = path.p, pattern = "csv$", recursive = TRUE))
# omit metadata from files to be read in as data
file.names <- file.names[!grepl("metadata", file.names)]   

# Load Surface area
SA <- read.csv(file="C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/Surface_Area/Surface_Area.csv", header=T)

# Load sample metadata
Sample.Info <- read.csv(file="C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/PR_use/20240810_11/pr_sample_metadata.csv", header=T) #read in sample.info data

# Load run metadata
run.info <- read_csv(file = "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/physiology/PR_use/20240810_11/pr_run_metadata.csv")
```

#Formating dataset for script
```{r, warning = FALSE}
# Join run and sample metadta first by sample id

colnames(SA)[colnames(SA) == "Colony_id"] <- "sample.id" #changing the column name in SA to match the sample metadata column name
metadata <- full_join(Sample.Info, run.info)%>%
  mutate(Date = as_date(as.character(Date), format = "%Y%m%d", tz = ""))


# Select only certain columnns
metadata <- metadata %>%
  select(sample.id, Run, Treatment, Species, Chamber.Vol.L, Date, Start.time, Stop.time, Light_Percent, Temp.Cat)

# Read in all data files
#edit sample.id name to remove .csv and the date from files
df <- tibble(file.name = file.names) %>%
  mutate(sample.id = gsub("_.*", "", file.name))%>%  # Get sample from filename and select for info before the first _
  mutate(sample.id = gsub(".csv", "", sample.id),  #remove .csv from sample.id
          info = map(sample.id, ~filter(metadata, sample.id == .)),           # Get associated sample info
         data0 = map(file.name, ~read_csv(file.path(path.p, .), skip = 1, col_types = cols(.default = "d", Time = "t"))))   # Get associated O2 data

# Select only Time, Value, and Temp columns from O2 data
df <- df %>%
  mutate(data0 = map(data0, ~select(., Time, Value, Temp)))%>%
  mutate(data0 = map(data0, ~(.x %>% filter(complete.cases(.))))) #remove NAs 

```

## Use the time breaks in the sample info to link O2 data with light levels
```{r, warning = FALSE}
df <- df %>%
  mutate(intervals = map2(data0, info, function(.x, .y) {
    split(.x, f = cut(as.numeric(.x$Time), breaks = as.numeric(c(.y$Start.time, last(.y$Stop.time))),
                      labels = as.character(.y$Light_Percent)))})) %>%
  mutate(data = map(intervals, ~ unnest(tibble(.), .id = "Light_Percent")))

## 'data' now contains the O2 data with the corresponding light level as another column
## Example of what 'data' for each sample looks like:
# df$data[[1]]

```


## Thin data
```{r, fig.height = 8, fig.width = 8}
# Set thinning parameter
thin_par <- 20

# Thin data for all samples
df <- df %>%
  mutate(thin_data = map(data, ~ slice(., seq(1, nrow(.), thin_par))))

# Create plots for full dataset and thinned data
df <- df %>%
  mutate(data_plot = map2(data, sample.id, ~ ggplot(.x, aes(x = Time, y = Value)) + 
                            facet_wrap(~ as.numeric(Light_Percent), scales = "free") +
                            geom_point() +
                            labs(title = .y)),
    thin_data_plot = map2(thin_data, sample.id, ~ ggplot(.x, aes(x = Time, y = Value)) + 
                            facet_wrap(~ as.numeric(Light_Percent), scales = "free") +
                            geom_point() +
                            labs(title = .y)))

# Example of plots
cowplot::plot_grid(df$data_plot[[1]], df$thin_data_plot[[1]], nrow = 2,
                   labels = c("Example plot: all data", "Example plot: thinned data"))



```

## The full or thinned data plot for any sample can be accessed like this:
```{r, warning = FALSE}
df %>%
  filter(sample.id == "MD-1-12") %>%
  pull(thin_data_plot)
```

# Fit regressions to each interval for each sample
```{r, warning = FALSE}
# Define function for fitting LoLinR regressions to be applied to all intervals for all samples
fit_reg <- function(df) {
  rankLocReg(xall = as.numeric(df$Time), yall = df$Value, 
             alpha = 0.2, method = "pc", verbose = FALSE)
}

# Setup for parallel processing
plan(multisession)

# Map LoLinR function onto all intervals of each sample's thinned dataset
df <- df %>%
  mutate(regs = furrr::future_map(thin_data, function(.) {       # future_map executes function in parallel
    group_by(., Light_Percent) %>%
    do(rankLcRg = fit_reg(.))
  }))

## Now 'regs' contains the fitted local regressions for each interval of each sample's thinned dataset

# Define function to pull out and plot regression diagnostics
plot_rankLcRg <- function(sample.id, interval_number) {
  df %>%
    filter(sample.id == sample.id) %>%
    pluck("regs", 1, "rankLcRg", interval_number) %>%
    plot()
}
```
## Extract slope of best regression for each interval for each sample
```{r}
df.out <- df %>% 
  unnest(regs) %>%
  mutate(micromol.L.s = map_dbl(rankLcRg, ~ pluck(., "allRegs", "b1", 1)))
```

## Adjust by chamber volume and normalize to surface area
```{r}
### Merge rates with sample info
pr <- left_join(
  select(df.out, sample.id, Light_Percent, micromol.L.s),
  distinct(metadata, sample.id, Treatment, Run, Chamber.Vol.L, Date)
)

## Correct for chamber volume and blanks
pr <- pr %>%
  mutate(micromol.s = micromol.L.s * Chamber.Vol.L)

# Get blank values -- average for each run and light value in case multiple blanks
blanks <- pr %>%
  filter(grepl("BLANK", sample.id)) %>%
  group_by(Run, Light_Percent) %>%
  summarise(micromol.s.blank = mean(micromol.s))

# Join blank values with rest of data and subtract values from samples for same run and light value
pr <- left_join(pr, blanks) %>%
  mutate(micromol.s.adj = micromol.s - micromol.s.blank) %>%
  # After correcting for blank values, remove blanks from data
  filter(!grepl("BLANK", sample.id))

# Join surface area with rest of data
pr <- left_join(pr, select(SA, sample.id, surface.area.cm2))

# Normalize rates by surface area
pr <- pr %>%
  mutate(micromol.cm2.s = micromol.s.adj / surface.area.cm2,
         micromol.cm2.h = micromol.cm2.s * 3600)

Photo <- pr %>%
  filter(Light_Percent ==331)%>% 
  separate(sample.id, c("Species","Genotype","Rep"), sep = "-", remove=FALSE)

Resp <- pr %>%
  filter(Light_Percent ==0)%>% 
  separate(sample.id, c("Species","Genotype","Rep"), sep = "-", remove=FALSE)

Photo <- na.omit(Photo) %>% select("sample.id", micromol.cm2.h, Treatment, Genotype, Species)
#colnames(Photo)[1] <- "sample.id"
Photo$Timepoint <- "Week5"
Photo$measure <- "Photo"
Resp <- na.omit(Resp) %>% select("sample.id", micromol.cm2.h, Treatment, Genotype, Species)
#colnames(Resp)[1] <- "sample.id"
Resp$Timepoint <- "Week5"
Resp$measure <- "Resp"

write.csv(Photo, file="C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Photo_Rawextracted_rates_Week5.csv")
write.csv(Resp, file="C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Resp_Rawextracted_rates_Week5.csv")
```

```{r}

df_summary_Resp <- Resp %>%
  group_by(Genotype, Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))

df_mean_Resp <- Resp %>%
  group_by(Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))


# Visualize data
pdf("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Raw_Resp_Week5.pdf", width=6, height=6)
r_plot<-df_summary_Resp %>%
    ggplot(., aes(x = Timepoint, y = -mean_micromol_cm2_h, colour=Treatment)) +
    geom_point(aes(fill=Treatment, group=Genotype, shape=Genotype), size=2) + 
    geom_point(data = df_mean_Resp, aes(x = Timepoint, y = -mean_micromol_cm2_h, colour=Treatment), size = 4) + 
    xlab("Timepoint") +
    facet_wrap(vars(Species))+
    scale_fill_manual(name="Treatment", values=c("gray","red","blue"))+
    scale_color_manual(name="Treatment", values=c("gray","red","blue"))+
    ylab(expression(bold(paste("Resp rate (µmol ", O[2], " cm"^-2, "min"^-1, ")")))) +
    #ylim(0,3)+
    theme_classic() + 
    theme(axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)); r_plot
if (dev.cur() > 1) {
  dev.off()
}

df_summary_Photo <- Photo %>%
  group_by(Genotype, Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))

df_mean_Photo <- Photo %>%
  group_by(Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))


# Visualize data
pdf("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Raw_Photo_Week5.pdf", width=6, height=6)
r_plot<-df_summary_Photo %>%
    ggplot(., aes(x = Timepoint, y = mean_micromol_cm2_h, colour=Treatment)) +
    geom_point(aes(fill=Treatment, group=Genotype, shape=Genotype), size=2) + 
    geom_point(data = df_mean_Photo, aes(x = Timepoint, y = mean_micromol_cm2_h, colour=Treatment), size = 4) + 
    xlab("Timepoint") +
    facet_wrap(vars(Species))+
    scale_fill_manual(name="Treatment", values=c("gray","red","blue"))+
    scale_color_manual(name="Treatment", values=c("gray","red","blue"))+
    ylab(expression(bold(paste("Photo rate (µmol ", O[2], " cm"^-2, "min"^-1, ")")))) +
    #ylim(0,3)+
    theme_classic() + 
    theme(axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12)); r_plot
if (dev.cur() > 1) {
  dev.off()
}

```
COMBINED PLOTS

Photosynthesis
```{r}
Pho.WK1 <- read.csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Photo_Rawextracted_rates_Week1.csv")
Pho.WK2 <- read.csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Photo_Rawextracted_rates_Week2.csv")
Pho.WK3 <- read.csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Photo_Rawextracted_rates_Week3.csv")
Pho.WK4 <- read.csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Photo_Rawextracted_rates_Week4.csv")
Pho.WK5 <- read.csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Photo_Rawextracted_rates_Week5.csv")

Pho <- rbind(Pho.WK1,Pho.WK3,Pho.WK4,Pho.WK5)
Pho$Genotype <- as.factor(Pho$Genotype)
df_summary_Pho <- Pho %>%
  group_by(Genotype, Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))
df_mean_Pho <- Pho %>%
  group_by(Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))

Pho <- Pho %>%
  select(sample.id, micromol.cm2.h, Species, Treatment, Timepoint) %>%
  mutate(Timepoint = case_when( 
    Timepoint == "Week1" ~ "TP1",
    Timepoint == "Week3" ~ "TP2",
    Timepoint == "Week4" ~ "TP3",
    Timepoint == "Week5" ~ "TP4",
    TRUE ~ NA_character_
  ))
frag_meta <- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/Coral_Fragment_Metadata.csv") %>%
  mutate(Sample.ID = gsub("_", "-", Sample.ID))

frag_meta <- frag_meta%>%
  select(Sample.ID, Inital.Treament, Secondary.Treatment, Treatment.Tank)%>%
  rename(Inital.Treatment = Inital.Treament)%>%
  rename(sample.id = Sample.ID)

Photo <- left_join (Pho, frag_meta, by = "sample.id")

Photo$new_timepoint <- ifelse(Photo$Timepoint == "TP4", paste0(Photo$Timepoint, "_", Photo$Secondary.Treatment), Photo$Timepoint)

Photo <- Photo %>%
  mutate(Species = case_when( 
    Species == "DL" ~ "Diploria labyrinthiformis",
    Species == "MD" ~ "Madracis decactis",
    TRUE ~ NA_character_
  ))
Photo <- Photo[-18, ]
```

Removing Outliers
```{r}
# Calculate the IQR
Q1 <- quantile(Photo$micromol.cm2.h, 0.25)
Q3 <- quantile(Photo$micromol.cm2.h, 0.75)
IQR <- Q3 - Q1

# Identify outliers
outliers <- Photo$micromol.cm2.h[(Photo$micromol.cm2.h < (Q1 - 1.5 * IQR)) | (Photo$micromol.cm2.h > (Q3 + 1.5 * IQR))]
print(outliers)

Photo <- Photo %>%
  filter(!(micromol.cm2.h %in% outliers))

#Removing Acclimation 
Photo2 <- Photo %>%
  filter(new_timepoint != "TP1")
```


```{r}

new_labels <- c("TP2" = "TP2","TP3" = "TP3", "TP4_Control" = "TP4 Control",
  "TP4_High_Temp" = "TP4 High Heat")

# Generate the plot OF chla
ggplot(Photo2, aes(x = new_timepoint, y = micromol.cm2.h, color = Inital.Treatment)) +
  geom_point(width = 0.1, position = position_jitterdodge(0.1), alpha = 0.1) +
  stat_summary(fun = mean, geom = "point", position = position_dodge(0.8)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5, position = position_dodge(0.8)) +
  scale_fill_manual(name = "Inital.Treatment", values = c("blue", "darkred", "purple"))+
  scale_color_manual(label = c("Control", "High Heat", "Low pH"), name = "Inital Treatment", values = c("blue", "darkred", "purple"))+
  scale_x_discrete(labels = new_labels) +
# facet_wrap(~ Species, nrow = 2) +
  facet_wrap(~ Species, nrow = 2) +
  labs(x = "Timepoint")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 15, hjust = 0.5, vjust = 0.6))+
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 1),
    strip.background = element_rect(colour = "black", fill = "white"),
    panel.spacing = unit(1, "lines"),  # Add spacing between facets
    panel.grid.major = element_blank(), # Remove major grid lines
    panel.grid.minor = element_blank(), # Remove minor grid lines
    strip.text = element_text(face = "bold")  # Make facet labels bold
  )+ 
  theme(
    strip.placement = "outside", # Position facet labels outside the panels
    strip.background = element_rect(colour = "black", fill = "white") # Borders around facet labels
  )
```
# Visualize data
pdf("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/All_Photo.pdf", width=6, height=6)

ggplot(df_summary_Pho, aes(x = Timepoint, y = mean_micromol_cm2_h, colour=Treatment)) +
  geom_boxplot()+
  facet_wrap(~Species)+
  theme_classic()

  ggplot(df_summary_Pho, aes(x = Timepoint, y = mean_micromol_cm2_h, colour=Treatment)) +
    geom_point(aes(fill=Treatment, group=Genotype, shape=Genotype), size=2) + 
    geom_point(data = df_mean_Pho, aes(x = Timepoint, y = mean_micromol_cm2_h, colour=Treatment), size = 4) + 
    xlab("Timepoint") +
    facet_wrap(vars(Species))+
    scale_fill_manual(name="Treatment", values=c("gray","red","blue","purple"))+
    scale_color_manual(name="Treatment", values=c("gray","red","blue","purple"))+
    ylab(expression(bold(paste("Photo rate (µmol ", O[2], " cm"^-2, "min"^-1, ")")))) +
    ylim(0,1.5)+
    theme_classic() + 
    theme(axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12))
dev.off()
  
```

#Respiration

```{r}
Res.WK1 <- read.csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Resp_Rawextracted_rates_Week1.csv")
Res.WK2 <- read.csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Resp_Rawextracted_rates_Week2.csv")
Res.WK3 <- read.csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Resp_Rawextracted_rates_Week3.csv")
Res.WK4 <- read.csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Resp_Rawextracted_rates_Week4.csv")
Res.WK5 <- read.csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/physiology/PR/Resp_Rawextracted_rates_Week5.csv")

Res <- rbind(Res_WK1,Res_WK3,Res_WK4,Res_WK5)
Res$Genotype <- as.factor(Res$Genotype)
df_summary_Res <- Res %>%
  group_by(Genotype, Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))
df_mean_Res <- Res %>%
  group_by(Treatment, Species, Timepoint) %>%
  summarise(mean_micromol_cm2_h = mean(micromol.cm2.h, na.rm = TRUE))
```

```{r}
Res <- Res %>%
  select(sample.id, micromol.cm2.h, Species, Treatment, Timepoint) %>%
  mutate(Timepoint = case_when( 
    Timepoint == "Week1" ~ "TP1",
    Timepoint == "Week3" ~ "TP2",
    Timepoint == "Week4" ~ "TP3",
    Timepoint == "Week5" ~ "TP4",
    TRUE ~ NA_character_
  ))
frag_meta <- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/Coral_Fragment_Metadata.csv") %>%
  mutate(Sample.ID = gsub("_", "-", Sample.ID))

frag_meta <- frag_meta%>%
  select(Sample.ID, Inital.Treament, Secondary.Treatment, Treatment.Tank)%>%
  rename(Inital.Treatment = Inital.Treament)%>%
  rename(sample.id = Sample.ID)

Resp <- left_join (Res, frag_meta, by = "sample.id")

Resp$new_timepoint <- ifelse(Resp$Timepoint == "TP4", paste0(Resp$Timepoint, "_", Resp$Secondary.Treatment), Resp$Timepoint)

Resp <- Resp %>%
  mutate(Species = case_when( 
    Species == "DL" ~ "Diploria labyrinthiformis",
    Species == "MD" ~ "Madracis decactis",
    TRUE ~ NA_character_
  ))
Resp <- Resp[-18, ]
```

Removing Outliers
```{r}
# Calculate the IQR
Q1 <- quantile(Resp$micromol.cm2.h, 0.25)
Q3 <- quantile(Resp$micromol.cm2.h, 0.75)
IQR <- Q3 - Q1

# Identify outliers
outliers <- Resp$micromol.cm2.h[(Resp$micromol.cm2.h < (Q1 - 1.5 * IQR)) | (Resp$micromol.cm2.h > (Q3 + 1.5 * IQR))]
print(outliers)

Resp2 <- Resp %>%
  filter(!(micromol.cm2.h %in% outliers))

#Removing Acclimation 
Resp3 <- Resp2 %>%
  filter(new_timepoint != "TP1")
```

```{r}

new_labels <- c("TP2" = "TP2","TP3" = "TP3", "TP4_Control" = "TP4 Control",
  "TP4_High_Temp" = "TP4 High Heat")

# Generate the plot OF chla
ggplot(Resp3, aes(x = new_timepoint, y = micromol.cm2.h, color = Inital.Treatment)) +
  geom_point(width = 0.1, position = position_jitterdodge(0.1), alpha = 0.1) +
  stat_summary(fun = mean, geom = "point", position = position_dodge(0.8)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5, position = position_dodge(0.8)) +
  scale_fill_manual(name = "Inital.Treatment", values = c("blue", "darkred", "purple"))+
  scale_color_manual(label = c("Control", "High Heat", "Low pH"), name = "Inital Treatment", values = c("blue", "darkred", "purple"))+
  scale_x_discrete(labels = new_labels) +
# facet_wrap(~ Species, nrow = 2) +
  facet_wrap(~ Species, nrow = 2) +
  labs(x = "Timepoint")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 15, hjust = 0.5, vjust = 0.6))+
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 1),
    strip.background = element_rect(colour = "black", fill = "white"),
    panel.spacing = unit(1, "lines"),  # Add spacing between facets
    panel.grid.major = element_blank(), # Remove major grid lines
    panel.grid.minor = element_blank(), # Remove minor grid lines
    strip.text = element_text(face = "bold")  # Make facet labels bold
  )+ 
  theme(
    strip.placement = "outside", # Position facet labels outside the panels
    strip.background = element_rect(colour = "black", fill = "white") # Borders around facet labels
  )
```

```{r}
# Visualize data
pdf("output/All_Respo.pdf", width=6, height=6)
df_summary_Res

ggplot(df_summary_Res, aes(x = Timepoint, y = mean_micromol_cm2_h, colour=Treatment)) +
  geom_boxplot()+
  facet_wrap(~Species)+
  theme_classic()


    ggplot(df_summary_Res, aes(x = Timepoint, y = -mean_micromol_cm2_h, colour=Treatment)) +
    geom_point(aes(fill=Treatment, group=Genotype, shape=Genotype), size=2) + 
    geom_point(data = df_mean_Res, aes(x = Timepoint, y = -mean_micromol_cm2_h, colour=Treatment), size = 4) + 
    xlab("Timepoint") +
    facet_wrap(vars(Species))+
    scale_fill_manual(name="Treatment", values=c("gray","red","blue","purple"))+
    scale_color_manual(name="Treatment", values=c("gray","red","blue","purple"))+
    ylab(expression(bold(paste("Resp rate (µmol ", O[2], " cm"^-2, "min"^-1, ")")))) +
    ylim(0,1.0)+
    theme_classic() + 
    theme(axis.title=element_text(face="bold", size=16),
      axis.text=element_text(size=12, color="black"), 
      legend.title=element_text(face="bold", size=14), 
      legend.text=element_text(size=12))
    
    
dev.off()
```


#Statistics

#Do stats on photo by treatment, species and date 
```{r}

TP2<- chl2%>%
  filter(timepoint != "TP0",
         timepoint !="TP1",
         timepoint !="TP3",
         timepoint !="TP4")

TP3<- chl2%>%
  filter(timepoint != "TP0",
         timepoint !="TP1",
         timepoint !="TP2",
         timepoint !="TP4")

TP4<- chl2%>%
  filter(timepoint != "TP0",
         timepoint !="TP1",
         timepoint !="TP2",
         timepoint !="TP3")
model <- aov(micromol.cm2.h ~ Treatment*Species*Timepoint, data = Resp3)
summary(model)
TukeyHSD(model)
#Shows significant differences between species, treatment, treatment-species, date-treatement and between date-species-treatment. 
to 
install.packages("emmeans")
install.packages("multcomp")
library(emmeans)
library(multcomp)

# View post-hoc differences 
emm = emmeans(model, ~ Treatment*Species*Timepoint)
cld(emm, Letters=c(LETTERS)) #letter display
pairs(emm)

res.aov <- aov(chla.ug.cm2 ~ Treatment*Species, data = chl1)
summary(res.aov)
TukeyHSD(res.aov)

```


import data Plot
```{r}

Photo <- rbind(Photo2,Photo3,Photo4,Photo5)
Photo <- Photo %>% rename(Sample.ID = sample.id)

use<-full_join(Photo, data)

Photo <- Photo[, -c(1, 4,5,8)]

class(Photo)
class(data)

data <- data[, -c(1:4,6,9,10,11)]

use <- use %>%
  dplyr::mutate(Treatmentstage = case_when(
    Timepoint == "Week4" ~ "Recovery",
    Timepoint == "Week5" ~ "Stress",
    Timepoint == "Week3" ~ "Priming",
    Timepoint == "Week2" ~ "Priming",
    TRUE ~ as.character(Timepoint)
  ))
  
df_summary <- use %>%
  group_by(Species, Treatmentstage, Inital.Treament,Secondary.Treatment) %>%
  summarize(
    photo_mean = mean(micromol.cm2.h),
    photo_se = sd(micromol.cm2.h) / sqrt(n()))
```


