---
title: "ENCORE POWER ANALYSIS"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This power analysis will determine the number of random fragmented corals to be sampled non destrutivley. With the back up fragments the current total is 508 fragments. 
Non destructive sampling includes fv/fm and respiration. 

Question - How do enviromental priming aid in coral adaptation mechanisms
Three Treatments - Control, Temperature and pH
We want to detect a 100%
I want to be 80% sure of picking up an effect

### two sample t test
Comparing the means of two independent groups (e.g., treatment vs. control group).
Example: You’re comparing the effectiveness of two different stress induced groups by analyzing performance  before and after exposure

### paried t test
This test is used when you have paired observations (e.g., before-and-after measurements on the same subjects).
Example: You’re assessing whether environmental priming aids in resilience by comparing pre and post measurments for the same individuals.

```{r}
library("pwr")
pwr.t.test(n =  , d = 1, sig.level =0.01 , power =0.8 , type = c("two.sample", "one.sample", "paried" ))
```

```{r}
library(pwr)

## Effect size 
# Define parameters
effect_size <- 0.2  
alpha <- 0.05
power <- 0.80
num_groups <- 6

# Calculate sample size per group
pwr.t.test(d = effect_size, power = power, sig.level = alpha, type = "two.sample")
#n= n = 5.089995


## Sample size of 5 
# Define parameters 
sample_size <- 5  
alpha <- 0.05
power <- 0.80
num_groups <- 6

# Calculate sample size 
pwr.t.test(n = sample_size, power = power, sig.level = alpha, type = "two.sample")
#effectsize = d =2.024439

# Sample size of 10 
sample_size <- 10  
# Calculate effect size 
pwr.t.test(n = sample_size, power = power, sig.level = alpha, type = "two.sample")
#effectsize = d = 1.324947

# Sample size of  
sample_size <- 21  
# Calculate effect size 
pwr.t.test(n = sample_size, power = power, sig.level = alpha, type = "two.sample")
#effectsize = d = d = 0.88605

# Sample size of  
sample_size <- 26  
# Calculate effect size 
pwr.t.test(n = sample_size, power = power, sig.level = alpha, type = "two.sample")
#d=0.0.7923522

```

