---
title: "Total Alkalinity with Parsing"
created by: Nyssa Silbiger 03/28/2014 and edited over the years by members of the Putnam Lab
date: "2024-10-08"
output: html_document
---

#This Script, calculates the total alkalinity using potentiometric titrations. Uses a for loop to read in data exported as a titration file and calculate Total alkalinity and at the end it exports your data as a .csv file.

#### Files needed ######
# 1. pHCalibration.csv in your "Data" folder
#Inside the Data folder You must have a subfolder for each data set. In each subfolder there is
# 2. the mass file for your run  
# 3. a subfolder named "TodaysDate" (where all of your titration files are) directly exported from LabX.



```{r setup, include=FALSE}

rm(list=ls()) # sweep environment
setwd("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/Titrator")


#NOTE: newer versions of the seacarb package have a hard time recognizing the "at" function. You need to check that the version installed is version 3.2, not version 3.3.1 which is the newest version of the "seacarb"

#packageurl <- "https://cran.r-project.org/src/contrib/Archive/seacarb/seacarb_3.2.tar.gz"
#install.packages(packageurl, repos=NULL, type="source")

library(seacarb) #used to calculate TA
library(tidyverse)
```


#THE DATA AND NAME OF FOLDER SHOULD BE CHANGED EACH TIME THE SCRIPT IS USED

```{r setup, echo==FALSE}
massfile<-"Mass_20241031_Samples.csv" # name of your file with masses
titrationfile<-'20241031_ENCORE_PRIMING_Samples.csv'# name of the last titration file run
date<-'20241031' #date that data was run
path<-"C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/Titrator/20241031" #the location of all your titration files, your folder of the day!

```

#LOAD MASS DATA

```{r setup, echo=FALSE}
#Load Mass Data
Mass<-read.csv(file.path(path,massfile), header=T, sep=",", na.string="NA", as.is=T) 
```


#### pH Calibration #####
```{r setup, echo=FALSE}

pHCal<-read.csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/Titrator/pHCalibration.csv") # read in the pH Calibration file

#select the calibration for the correct date
pHData<-pHCal[pHCal$Date==date,]

#calculate pH 3 and 3.5 based on the slope and intercept from pH 4, 7, and 10 calibration
mod.pH<-lm(c(pHData$pH4, pHData$pH7, pHData$pH10)~c(4,7,10)) # linear model

# print a plot of the relationship between pH and mV
png(paste0(path,"/",Sys.Date(),'pHmvplot.png'), height = 400, width = 400)

# Initialize the plot
plot(c(4,7,10), c(pHData$pH4, pHData$pH7, pHData$pH10), xlab = 'pH', ylab = 'mv')

# Add lines to the existing plot
lines(c(4,7,10), predict(mod.pH))

R2<-summary(mod.pH)$r.squared
legend('topright', legend = bquote(R^2 == .(format(R2, digits = 3))), bty='n')
dev.off()

# Select the mV for pH=3 and pH=3.5 based on your probe calibration
pH35<-mod.pH$coefficients[1]+mod.pH$coefficients[2]*3.5
pH3<-mod.pH$coefficients[1]+mod.pH$coefficients[2]*3
```

```{r setup, echo=FALSE}
##### titration###########

#create an empty matrix to put the TA values in

nrows<-nrow(Mass) #need file length/number of rows


TA <- data.frame(matrix(nrow = nrows, ncol = 5)) # changes from 4 columns to 5

rownames(TA)<-Mass$Sample.ID1[1:nrows]

colnames(TA)<-c("Date","SampleID",'TA','Mass','Salinity') # added a date and tank column

# changed Sample.ID1 to SampleID in the TA data frame only

# read in the mega concatenated titration results file

filename<-file.path(path,titrationfile)

AllData<-read.csv(filename, sep=",", na.string="NA", as.is=T, skip=4)[ ,1:5]

#Identifies rows starting with scope in column 1 of the titration file
sample_name_positions <- c(1,grep("^Scope", AllData[,1]), nrow(AllData))

## parse through all the data in the one file ###
sample_names<-Mass$Sample.ID1

# create a list with all the sample IDs
sample_names_list <- list()
for (item in 1:length(sample_names)){
  sample_names_list[[item]] <- sample_names[item]
}

# fill the list with the data from each sample
for (i in 1:nrows){
  sample_names_list[[i]]<-data.frame(AllData[sample_name_positions[i]:sample_name_positions[i+1],])
  colnames(sample_names_list[[i]])<-c("Volume","Time","mV","Temperature","dV/dt")
}

```

```{r setup, echo=FALSE}

for(i in 1:nrows) {
  
  #  Data<-read.csv(file.names[i], header=F, sep=",", na.string="NA",as.is=T, skip=10)[ ,1:5] 
  
  # colnames(Data) <-  c("Volume","Time",	"mV",	"Temperature",	"dV/dt")
  
  Data<-sample_names_list[[i]]
  
  # everything was brought in as a character because of the second line, converts back to numeric
  
  Data$mV<-suppressWarnings(as.numeric(Data$mV)) ## suppress the warnings since NA will be produced
  
  Data$Temperature<-suppressWarnings(as.numeric(Data$Temperature))
  
  Data$Volume<-suppressWarnings(as.numeric(Data$Volume))
  
  #name of the file without .csv
  
  #name<-unlist(strsplit(file.names[i], split='.', fixed=TRUE))[1]
  
  name<-sample_names[i]
  
  
  
  #calculates the index of values between pH 2 and 3.5 
  
  mV<-which(Data$mV<pH3 & Data$mV>pH35) 
 
  
  #CHANGE ONLY WHEN NEW BOTTLE OF ACID IS USED----------------------------------
  
  #density of your titrant: change every time acid is changed
  
  #Florence Fields updated script on 20241008 with the current batch opened 20240613 #A24
 
  d <- (-0.00000410*mean(Data$Temperature[mV], na.rm=T)^2-0.0001069*mean(Data$Temperature[mV], na.rm=T)+1.02881) #20240613 Batch A24 FF
  
  
  c<-0.099922 ##Batch A24 currently in use since 20240613 -FF
  
  
  #------------------------------------------------------------------------------
  
  
  
  #Salinity of your samples, set salinity for all titrations using a junk or CRM (single value)
  s<-Mass[Mass$Sample.ID1==name,3]
  #s<-Mass[name,2]
  
  #mass of sample in g: changed with every sample
  #mass<-Mass[name,1]
  mass<-Mass[Mass$Sample.ID1==name,2]
  #sample.index<-Mass[Mass$Sample.ID1==name,3]# this is the order that the sample was run
  #-------------------------------------------------------------------
  #Calculate TA
  
  #at function is based on code in seacarb package by Steve Comeau, Heloise Lavigne and Jean-Pierre Gattuso
  
  
  TA[i,1]<-date #exports the date into output file, column 1; added by LZ 20220720
  TA[i,2]<-name #exports the sample ID into output file, column 2
  TA[i,3]<-1000000*at(S=s,T=mean(Data$Temperature[mV], na.rm=T), C=c, d=d, pHTris=NULL, ETris=NULL, weight=mass, E=Data$mV[mV], volume=Data$Volume[mV])
  TA[i,4]<-mass #exports the mass into the TA output file, column 4
  TA[i,5]<-s #exports the salinity column into the output file, column 5
}
```

```{r setup, echo=FALSE}
TA[,3:4]<-sapply(TA[,3:4], as.numeric) # make sure the appropriate columns are numeric


#Renaming Junk_3 to fit into the data set
TA$SampleID[TA$SampleID == "Junk_3"] <- "Junk_1"

# Use separate() to split the 'Name' column into 'FirstName' and 'LastName'
TA <- TA %>% separate(SampleID, into = c("DateSampled", "Tank"), sep = "_")

print(TA)


TA$PrimingTreatment <- ifelse(TA$Tank == 7, "Heat", 
                              ifelse(TA$Tank == 5, "Control", 
                                     ifelse(TA$Tank == 2, "pH", 
                                            ifelse(TA$Tank == 4, "Control", 
                                                   ifelse(TA$Tank == 1, "Junk",NA)))))

print(TA)

TA[,4:6]<-sapply(TA[,4:6], as.numeric) # make sure the appropriate columns are numeric

#Renaming S20240809, S20240726, S20240802, S20240809 to fit into the data set
TA$DateSampled[TA$DateSampled == "S20240809"] <- "20240809"
TA$DateSampled[TA$DateSampled == "S20240726"] <- "20240726"
TA$DateSampled[TA$DateSampled == "S20240802"] <- "20240802"
TA$DateSampled[TA$DateSampled == "S20240809"] <- "20240809"



#exports your data as a CSV file
write.table(TA,paste0(path,"/","TA_Output_",titrationfile),sep=",", row.names=FALSE)

#Removing values that were redone


#Cumulative TA
cumu.data <- read.csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/Titrator/Cumulative_TA_Output.csv", header=TRUE, sep=",")
update.data <- rbind(cumu.data, TA)

#check that your new data has been appended to the cumulative TA dataframe (added 20220623 by LZ)
tail(update.data)

getwd()
#export data as csv file
write.table(update.data,"C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/Titrator/Cumulative_TA_Output.csv",sep=",", row.names=FALSE)
```


Plot the TA date excluding vaules from above 2900 and below 2000               
```{r setup, echo=FALSE}

#Can either use update.data from the dataframe currently in the environment of import the dataset

Cumulative_TA_Output <- read_csv("C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/data/Titrator/Cumulative_TA_Output.csv")

library(dplyr)

# Remove rows where DateSampled is "junk"
cleaned_data <- Cumulative_TA_Output %>% 
  filter(DateSampled != "Junk")

#Create the plot
plot<-ggplot(cleaned_data, aes(DateSampled, TA, color=Tank)) +
  geom_point() +
  ylim(2000, 2800)+
  scale_color_manual(values = c("pink", "purple", "yellow", "green", "blue", "red", "black", "brown")) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme(legend.box.background = element_rect(color = "black"), 
        legend.box.margin = margin(t = 1, l = 1), 
        panel.border = element_rect(color = "black", fill = NA, size = 1))+
   facet_wrap(~PrimingTreatment,, scales= "free_y", nrow=4) 

# Save the plot
ggsave(plot = plot, path = "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/titrator", 
       filename = "New_TA_data_plot_by_treatment2.pdf")
```

```{r setup, echo=FALSE}

sorted_data <- cleaned_data %>% 
  arrange(DateSampled)
sorted_data <- sorted_data %>% arrange(Tank, DateSampled)
#sorted_data$DateSampled <- as.Date(sorted_data$DateSampled, format = "%Y-%m-%d")


#Create the plot
ggplot(sorted_data, aes(DateSampled, TA, color=Tank)) +
  geom_point()+geom_line()+
  ylim(2000, 2800) +
  scale_color_manual(values = c("pink", "purple", "yellow", "green", "blue", "red", "black", "brown")) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme(legend.box.background = element_rect(color = "black"), 
        legend.box.margin = margin(t = 1, l = 1), 
        panel.border = element_rect(color = "black", fill = NA, size = 1))+
  facet_wrap(~PrimingTreatment, scales= "free_y", ncol=4)

ggsave(plot = plot2, path = "C:/Users/flo_f/OneDrive - University of Rhode Island/GitHub/Coral_Priming_Experiments_Summer_2024/output/titrator", 
       filename = "New_TA_data_line_plot.pdf")
```
